---
title: "simplified_V1_playground"
author: "Richard Schweitzer"
date: "2023-05-26"
output: github_document
---

Libraries...

```{r}
library(assertthat)
library(data.table)
library(ggplot2)
library(scales)
library(viridis)
library(cowplot)
library(pracma)
library(mgcv)
library(torch)
```

Screen parameters and more
```{r}
# monitor parameters
monitor.dist <- 180 # distance to monitor in cm
monitor.width <- 200 # width of monitor in cm
monitor.resx <- 960 # horizontal pixels of monitor
monitor.resy <- 540 # vertical pixels of monitor
monitor.screen_center <- data.frame(x = monitor.resx/2, y = monitor.resy/2)
# old and wrong value (monitor.dist was set to 200 instead of 180) for pixels per dva on the screen:
monitor.screen_ppd_old <- 16.7569 
monitor.screen_ppd <- monitor.dist*tan(1*pi/180)/(monitor.width/monitor.resx) # correct
monitor.ppd_correction_factor <- monitor.screen_ppd / monitor.screen_ppd_old
# update stimulus parameters
sti.freq <- round(c(0.0298/2, 0.0298, 0.0298*2)*monitor.screen_ppd, 2)  # sti.freqp = 0.0298
sti.sd <- round(8.3778 / monitor.screen_ppd, 2) # sti.sdp = 8.3778
sti.sac_amp <- round(275.2756 / monitor.screen_ppd, 2) # dot.sac_length = 275.2756
```

Global modeling parameters
```{r}
scr.ppd <- monitor.screen_ppd
spatial_resolution = 0.05 # spatial resolution of processing function in dva
temporal_resolution = 1000/1440 # temporal resolution of processing function in milliseconds
(SFs = exp(linspace(log(0.03), log(3), 9)))
(Oris = seq(-pi/2, pi/2, length.out = 11)[-1])
Phases = c(0, pi/2)
```

Prepare list of Gabor filters and matching TRFs

```{r}
source("get_IRF_df.R")
irf_space <- get_IRF_df(SFs = SFs, irf_time_range = c(0, 300), 
                        temporal_resolution = temporal_resolution)
# sum for each SF
irf_space[ , .(irf_sum = sum(abs(irf))), by = .(SF)]
# plot the predictions
ggplot(data = irf_space, aes(x = irf_time, y = SF, fill = irf)) + 
  geom_raster(interpolate = TRUE) + 
  scale_y_log10() + 
  scale_fill_gradient2() + theme_minimal()
```

... and the matching Gabor filters:
```{r}
source("plot_heatmap.R")
source("get_gabor_field.R")
source("get_gabor_filter_bank.R")

# create the Gabor filters
gabor_list <- get_gabor_filter_bank(
  all_SF = SFs, 
  all_Ori = Oris, # positive means clockwise
  rf_width_override = (1/SFs) / 2, # cpd, set to NaN to let the function determine RF field size
  scr_ppd = scr.ppd, 
  ppd_scaler = 1,
  make_gaussian_aperture = TRUE)
# how many have we got?
length(gabor_list)
gabor_list[[1]][[3]]

# a sanity check
p_gabor_list_1 <- vector(mode = "list")
p_gabor_list_2 <- vector(mode = "list")
for (gi in 1:length(gabor_list)) {
  p_gabor_list_1[[gi]] <- plot_heatmap(mat = gabor_list[[gi]][[1]], do_print = FALSE) + 
    theme(legend.position = "none") + theme_nothing()
  p_gabor_list_2[[gi]] <- plot_heatmap(mat = gabor_list[[gi]][[2]], do_print = FALSE) + 
    theme(legend.position = "none") + theme_nothing()
}
# plot 1: zero phase
plot_grid(plotlist = p_gabor_list_1, nrow = length(SFs))
# plot 2: half phase
plot_grid(plotlist = p_gabor_list_2, nrow = length(SFs))
```

Look for a good normalization for the Gabor filter. 
```{r}
norm_resps <- NULL
for (gabor_i in 1:length(gabor_list)) {
  # these are the gabors
  this_gabor_zero <- gabor_list[[gabor_i]][[1]]
  this_gabor_half <- gabor_list[[gabor_i]][[2]]
  # these are the max-response patches
  sign_gabor_zero <- rnorm(length(this_gabor_half))/6+(this_gabor_zero)
  sign_gabor_half <- rnorm(length(this_gabor_half))/6+(this_gabor_half)
#  sign_gabor_zero <- sign(this_gabor_zero)
#  sign_gabor_half <- sign(this_gabor_half)
  # responses
  resp_zero <- sum(this_gabor_zero*sign_gabor_zero)
  resp_half <- sum(this_gabor_half*sign_gabor_half)
  # normalize?
  resp_zero_norm <- resp_zero / sum(abs(this_gabor_zero))
  resp_half_norm <- resp_half / sum(abs(this_gabor_half))
  # save
  norm_resps <- rbind(norm_resps, 
                      data.table(SF = as.numeric(gabor_list[[gabor_i]][[3]]['rf_freq_dva']), 
                                 Ori = as.numeric(gabor_list[[gabor_i]][[3]]['rf_ori']), 
                                 dim_x = dim(this_gabor_zero)[1], 
                                 dim_y = dim(this_gabor_half)[1], 
                                 resp_zero = resp_zero, 
                                 resp_half = resp_half, 
                                 resp_zero_norm = resp_zero_norm, 
                                 resp_half_norm = resp_half_norm))
}
norm_resps
# look at how responses scale with Gabor size
ggplot(data = norm_resps, aes(x = SF, y = sqrt(resp_zero^2 + resp_half^2), 
                              color = Ori, group = Ori )) + 
  geom_point() + geom_line() + 
  scale_x_log10() + #scale_y_log10() + 
  scale_color_gradient2()
# look at normalization success:
ggplot(data = norm_resps, aes(x = SF, y = sqrt(resp_zero_norm^2 + resp_half_norm^2), 
                              color = Ori, group = Ori )) + 
  geom_point() + geom_line() + 
  scale_x_log10() + #scale_y_log10() + 
  scale_color_gradient2()
```

Seems like dividing with the absolute sum is good enough. 

Now load a random stimulus from SDEC and create arbitrary trajectory
```{r}
source("load_noise_patch.R")
noise_patch <- load_noise_patch(stim_name = "stim_SDEC0201.csv")
plot_heatmap(noise_patch) # plot it
# create an arbitrary trajectory
x_stim_traj <- c(345.080798685704, 342.197939661274, 339.315080636844, 336.432221612414, 
                 333.549362587984, 330.666503563554, 327.783644539124, 324.900785514694, 
                 322.017926490264, 319.135067465834, 316.252208441404, 313.369349416974, 
                 310.486490392544, 307.603631368114, 304.720772343685, 301.837913319255, 
                 298.955054294825, 296.072195270395, 293.189336245965, 290.306477221535, 
                 287.423618197105, 284.540759172675, 281.657900148245, 278.775041123815, 
                 275.892182099385, 273.009323074955, 270.126464050525, 267.243605026095,
                 264.360746001665, 261.477886977235, 258.595027952805, 255.712168928375, 
                 252.829309903945, 249.946450879515, 247.063591855085, 244.180732830655) 
y_stim_traj <- rep(275, length.out = length(x_stim_traj)) + 1:length(x_stim_traj)
t_stim_traj <- seq(10, 10+(length(x_stim_traj)-1)*temporal_resolution, 
                   by = temporal_resolution)
assert_that(length(x_stim_traj)==length(t_stim_traj))
```

Construct the x-y-t matrix and set up convolutions

```{r}
# to debug:
# signal_x = x_stim_traj / scr.ppd
# signal_y = y_stim_traj / scr.ppd
# signal_t = t_stim_traj
# irf_df = irf_space
# stim_mat = noise_patch
# scr_ppd = scr.ppd

# source the two spatial convolution functions here:
source("standard_conv_with_torch_3d.R")
source("fft_conv_with_torch_3d.R")

# this is the V1 module
v1 <- function(stim_mat, # the noise patch as a matrix
               gabor_list, # a list of gabor filters created by 'get_gabor_filter_bank'
               irf_df, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
               scr_ppd, # pixels per degree to convert matrix sizes to cycles per degree
               signal_x, signal_y, signal_t, # properties of the signal (relative to saccade onset)
               output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
               signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
               no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
               use_half_on_GPU = FALSE, # if a GPU is available, use half-precision?
               spatial_resolution = 0.05, # spatial resolution of processing function in dva
               temporal_resolution = 1000/1440, # temporal resolution of processing function in milliseconds
               debug_mode = FALSE, # shows output at every step
               show_final_maps = TRUE # shows all resulting 2D maps at the end
) {
  require(data.table)
  require(assertthat)
  require(torch)
  require(pracma)
  require(tictoc)
  ## 0.1 quick checks
  # do the SFs of the supplied IRF and Gabor objects match?
  gabor_SFs <- sort(unique(as.numeric(unlist(lapply(X = gabor_list, FUN = function(x) { x[[3]]['rf_freq_dva'] })))))
  irf_SFs <- sort(unique(irf_df$SF))
  are_equal(gabor_SFs, irf_SFs)
  # have we loaded the convolution functions?
  assert_that(exists("standard_conv_with_torch_3d"))
  assert_that(exists("fft_conv_with_torch_3d"))
  assert_that(exists("plot_heatmap"))
  # is the IRF properly scaled in the temporal domain?
  irf_Time <- sort(unique(irf_df$irf_time))
  are_equal(length(unique(round(diff(irf_Time), 5))), 1)
  are_equal(unique(round(diff(irf_Time), 5)), round(temporal_resolution, 5))
  ## 0.2 compute properties of signal
  length_signal_x <- abs(max(signal_x)-min(signal_x)) # the spatial extent of the signal 
  length_signal_y <- abs(max(signal_y)-min(signal_y))
  length_signal_t <- abs(max(signal_t)-min(signal_t)) # the spatial extent of the signal 
  assert_that(length(signal_x)==length(signal_y))
  assert_that(length(signal_x)==length(signal_t))
  n_samples <- length(signal_t)
  signal_df <- data.frame(signal_x, signal_y, signal_t)
  signal_df$n_samples <- 1:n_samples
  if (debug_mode) {
    print("Input:")
    print(signal_df)
  }
  ## STEP 1: fill the position-over-time matrix
  # 1.1 get the dimensions for the cells of the matrix
  spatial_pad_size <- round(max(dim(stim_mat)) / scr_ppd)
  temporal_pad_size <- round(1*(max(irf_Time)-min(irf_Time)))
  # X
  if (!is.null(signal_x_range)) {
    mat_size_x = seq(signal_x_range[1]-spatial_pad_size, signal_x_range[2]+spatial_pad_size, 
                     by = spatial_resolution) # columns, x coordinate
  } else {
    mat_size_x = seq(min(signal_x)-spatial_pad_size, max(signal_x)+spatial_pad_size, 
                     by = spatial_resolution) # columns, x coordinate
  }
  # Y
  if (!is.null(signal_y_range)) {
    mat_size_y = seq(signal_y_range[1]-spatial_pad_size, signal_y_range[2]+spatial_pad_size, 
                     by = spatial_resolution) # rows, y coordinate
  } else {
    mat_size_y = seq(min(signal_y)-spatial_pad_size, max(signal_y)+spatial_pad_size, 
                     by = spatial_resolution) # rows, y coordinate
  }
  # T
  if (!is.null(signal_t_range)) { # compute the temporal scale based on signal_t_range
    mat_size_t = seq(signal_t_range[1], signal_t_range[2]+temporal_pad_size, 
                     by = temporal_resolution)
  } else { # compute the temporal scale based on signal_t
    if (min(signal_t)>=0) {
      min_t <- 0
    } else {
      min_t <- min(signal_t)
    }
    mat_size_t = seq(min_t, max(signal_t)+temporal_pad_size, 
                     by = temporal_resolution) # third dimension, t coordinate (0 was previously: min(signal_t)-round(length_signal_t))
  }
  # 1.3 create large matrix of zeros according to the specified resolution
  mat_over_t <- array(data = 0, dim = c(length(mat_size_y), length(mat_size_x), length(mat_size_t)), 
                      dimnames = list(mat_size_y, mat_size_x, mat_size_t) )
  if (debug_mode) {
    print(paste("dim(mat_over_t) =", paste0(dim(mat_over_t), collapse = ","), "[y,x,t]" ))
  }
  # 1.4 fill the matrix with the blob where the position is present in the signal
  p_stim_sequence_list <- vector(mode = "list", length = n_samples)
  for (sample_i in (1:n_samples)) {
    put_sample_x_here <- which.min(abs(mat_size_x - signal_x[sample_i]))
    put_sample_y_here <- which.min(abs(mat_size_y - signal_y[sample_i]))
    put_sample_t_here <- which.min(abs(mat_size_t - signal_t[sample_i]))
    mat_over_t[seq(put_sample_y_here-round(nrow(stim_mat)/2), 
                   put_sample_y_here-round(nrow(stim_mat)/2)+nrow(stim_mat)-1, 
                   by = 1), 
               seq(put_sample_x_here-round(ncol(stim_mat)/2), 
                   put_sample_x_here-round(ncol(stim_mat)/2)+ncol(stim_mat)-1, 
                   by = 1), 
               put_sample_t_here] <- stim_mat
    #print(address(mat_over_t)) # this operation works in place
    if (debug_mode) {
      p_stim_now <- plot_heatmap(mat_over_t[ , , put_sample_t_here], do_print = FALSE)
      p_stim_sequence_list[[sample_i]] <- p_stim_now + 
        #ggtitle(paste("t =", signal_t[sample_i])) + 
        theme(legend.position = "none")
    }
  }
  # plot the sequence?
  if (debug_mode) {
    require(cowplot)
    p_stim_sequence <- plot_grid(plotlist = p_stim_sequence_list)
    print(p_stim_sequence)
  }
  last_signal_t_index <- put_sample_t_here # save the last indeces where we have a stimulus present
  
  ## STEP 2: Convolutions performed in loop over SF and Ori
  # is cuda available?
  if (cuda_is_available()) {
    if (debug_mode) {
      print(paste("CUDA is available! Version:", cuda_runtime_version()))
    }
    if (no_CUDA) {
      use_device <- torch_device('cpu')
    } else {
      use_device <- torch_device('cuda') # 'cuda'
    }
    # ... and what should be the preferred data type? (there are torch_half, torch_float, torch_double)
    if (use_half_on_GPU) {
      preferred_dtype <- torch_float16() # or half as mixed/half computing is optimized?
    } else {
      preferred_dtype <- torch_float32()
    }
  } else {
    use_device <- torch_device('cpu')
    preferred_dtype <- torch_float32()
  }
  # 0) reshape the mat_over_t to match the 4D requirements of the 2D convolution
  # transfer to GPU, if necessary:
  mat_over_t_gpu <- torch_tensor(mat_over_t, device = use_device, dtype = preferred_dtype)
  mat_over_t_gpu <- mat_over_t_gpu$permute(c(3, 1, 2))$unsqueeze(1)
  if (debug_mode) {
    print("Converted matrix to tensor and reshaped to dimensions:")
    print(mat_over_t_gpu$size())
  }
  # preallocate the final output matrix [filter, y, x]
  final_2d_over_filters <- array(data = NaN, 
                                 dim = c(length(gabor_list), dim(mat_over_t)[1], dim(mat_over_t)[2]))
  SFs_of_filters <- vector(mode = "numeric", length = length(gabor_list))
  Oris_of_filters <- vector(mode = "numeric", length = length(gabor_list))
  if (output_full_sequences) {
    final_3d_over_filters <- array(data = NaN, 
                                   dim = c(length(gabor_list), dim(mat_over_t)[1], dim(mat_over_t)[2], dim(mat_over_t)[3]))
    # [filter, y, x, t]
  } else {
    final_3d_over_filters <- NULL
  }
  # this is THE LOOP !
  for (gabor_list_i in 1:length(gabor_list)) {
    # which SF are we dealing with?
    SF_now <- gabor_list[[gabor_list_i]][[3]]['rf_freq_dva']
    Ori_now <- gabor_list[[gabor_list_i]][[3]]['rf_ori']
    if (debug_mode || show_final_maps) { 
      print(paste0("NEW Gabor Nr. ", gabor_list_i, ", SF=", round(SF_now, 2), ", Ori=", round(Ori_now, 2) )) 
    }
    # get corresponding IRF and normalize
    IRF_now <- irf_df[SF==SF_now, irf]
    IRF_now_time <- irf_df[SF==SF_now, irf_time]
    IRF_now <- IRF_now / sum(abs(IRF_now))
    # get Iris Groen's low-pass filter for the delayed-normalization model
    prm.tau2 <- 0.75 # see Figure 11 from "Temporal Dynamics of Neural Responses in Human Visual Cortex"
    IRF_lp = exp(-(IRF_now_time/1000)/prm.tau2)
    IRF_lp = IRF_lp / sum(abs(IRF_lp))
    # get Gabor filter kernels (1: zero, 2: half phase) and make sure they have odd dimensions
    spat_kernel_1 <- gabor_list[[gabor_list_i]][[1]]
    spat_kernel_2 <- gabor_list[[gabor_list_i]][[2]]
    assert_that(all(mod(dim(spat_kernel_1), 2) == 1), msg = "spat_kernel_1 does not have odd dimensions")
    assert_that(all(mod(dim(spat_kernel_2), 2) == 1), msg = "spat_kernel_2 does not have odd dimensions")
    # normalize
    spat_kernel_1 <- spat_kernel_1 / sum(abs(spat_kernel_1))
    spat_kernel_2 <- spat_kernel_2 / sum(abs(spat_kernel_2))
    ## 1) determine how convolution should be performed - and perform it
    # compute number of operations. If there are more than 10^8.5 in the spatial domain, use FFT-based
    number_of_spatial_operations <- prod(dim(spat_kernel_1)) * 
      mat_over_t_gpu$size(3) * mat_over_t_gpu$size(4)
    if ( !cuda_is_available() || number_of_spatial_operations > 10^8.5) {
      use_fft_based_2dconv <- TRUE
    } else {
      use_fft_based_2dconv <- FALSE
    }
    ## 2) run the spatial convolution
    # note that pytorch implements a cross-correlation, not a convolution. 
    # Thus, we'll have opposite results as when done in the fft domain
    if (use_fft_based_2dconv) { # fft-based convolution
      tic(msg = "fft-based 2D convolution")
      spatial_response_1 <- fft_conv_with_torch_3d(x = mat_over_t_gpu, K = spat_kernel_1, use_torch = TRUE,
                                                   no_CUDA = no_CUDA, use_half_on_GPU = use_half_on_GPU)
      spatial_response_2 <- fft_conv_with_torch_3d(x = mat_over_t_gpu, K = spat_kernel_2, use_torch = TRUE,
                                                   no_CUDA = no_CUDA, use_half_on_GPU = use_half_on_GPU)
      toc(quiet = !(debug_mode || show_final_maps))
    } else { # normal convolution
      tic(msg = "standard 2D convolution")
      spatial_response_1 <- standard_conv_with_torch_3d(x = mat_over_t_gpu, K = spat_kernel_1, 
                                                        no_CUDA = no_CUDA, use_half_on_GPU = use_half_on_GPU)
      spatial_response_2 <- standard_conv_with_torch_3d(x = mat_over_t_gpu, K = spat_kernel_2, 
                                                        no_CUDA = no_CUDA, use_half_on_GPU = use_half_on_GPU)
      toc(quiet = !(debug_mode || show_final_maps))
    }
    assert_that(all(mat_over_t_gpu$size()==spatial_response_1$size()))
    assert_that(all(mat_over_t_gpu$size()==spatial_response_2$size()))
    # ... diagnostic plots: what does the convolution look like?
    if (debug_mode) {
      print("Performed 2D spatial convolution, thereby achieving matrix with size:")
      print(paste(paste0(spatial_response_1$size(), collapse = ","), "and", paste0(spatial_response_2$size(), collapse = ",")))
      p_spatial_responses <- vector(mode = "list", length = 5)
      p_spatial_responses[[1]] <- plot_heatmap(as_array(mat_over_t_gpu[1, last_signal_t_index-3, , ]$cpu()), do_print = FALSE) + 
        ggtitle("Original stimulus")
      p_spatial_responses[[2]] <- plot_heatmap(rbind(spat_kernel_1, spat_kernel_2), do_print = FALSE) + 
        ggtitle("Spatial kernels (zero & half)")
      p_spatial_responses[[3]] <- plot_heatmap(as_array(spatial_response_1[1, last_signal_t_index-3, , ]$cpu()), do_print = FALSE) + 
        ggtitle(paste0("Spatial resp (zero), SF=", round(SF_now, 2), ", Ori=", round(Ori_now, 2) ))
      p_spatial_responses[[4]] <- plot_heatmap(as_array(spatial_response_2[1, last_signal_t_index-3, , ]$cpu()), do_print = FALSE) + 
        ggtitle(paste0("Spatial resp (half), SF=", round(SF_now, 2), ", Ori=", round(Ori_now, 2) ))
      p_spatial_responses[[5]] <- plot_heatmap(sqrt(as_array(spatial_response_1[1, last_signal_t_index-3, , ]$cpu())^2 + 
                                                    as_array(spatial_response_2[1, last_signal_t_index-3, , ]$cpu())^2), do_print = FALSE) + 
        ggtitle(paste0("Squared responses"))
      p_spatial_responses <- plot_grid(plotlist = p_spatial_responses, nrow = 3)
      print(p_spatial_responses)
    } else {
      rm(spat_kernel_1, spat_kernel_2)  # save memory!
    }
    ## 3) prepare the temporal response kernel (is has to have odd dimension - and the procedure below assures that)
    # spatial_response is now [1, t, y, x]
    resp_fun_gpu <- torch_tensor(data = IRF_now, device = use_device, dtype = preferred_dtype)
    resp_fun_gpu <- resp_fun_gpu$expand(c(spatial_response_1$size(4), 1, resp_fun_gpu$size())) # expand to match matrix dimensions (x dimension)
    resp_fun_gpu <- torch_cat(list(torch_zeros(c(resp_fun_gpu$size(1), resp_fun_gpu$size(2), resp_fun_gpu$size(3)-1), 
                                               device = use_device, dtype = preferred_dtype), # create the zero pad to center the TRF
                                   resp_fun_gpu), dim = 3)
    if (debug_mode) {
      print("Created temporal response tensor of dimensions for 1D convolution:")
      print(resp_fun_gpu$size())
    }
    ## 4) prepare and run the temporal convolution
    spatial_response_1 <- spatial_response_1$permute(c(1, 3, 4, 2))$squeeze() # new dim: y, x, t, as initially
    spatial_response_2 <- spatial_response_2$permute(c(1, 3, 4, 2))$squeeze()
    if (debug_mode) {
      print("Reshaped main tensor to dimensions to prepare for 1D temporal convolution:")
      print(spatial_response_1$size())
    }
    temporal_response_1 <- torch_conv1d(input = spatial_response_1, 
                                        weight = resp_fun_gpu$flip(3), 
                                        padding = resp_fun_gpu$size(3)%/%2, 
                                        groups = spatial_response_1$size(2) # is the x dimension
    )
    temporal_response_2 <- torch_conv1d(input = spatial_response_2, 
                                        weight = resp_fun_gpu$flip(3), 
                                        padding = resp_fun_gpu$size(3)%/%2, 
                                        groups = spatial_response_2$size(2)
    )
    # checks:
    are_equal(temporal_response_1$size(), temporal_response_2$size())
    are_equal(temporal_response_1$size(), spatial_response_1$size())
    are_equal(temporal_response_2$size(), spatial_response_2$size())
    if (debug_mode) {
      print("Performed convolution through time, resulting in matrix of dim:")
      print(temporal_response_1$size())
      # progress of response through time
      temporal_resp_plots <- vector(mode = "list", length = 16)
      for (t_index_now_i in 1:16) {
        t_index_now <- seq(1, temporal_response_1$size(3), length.out = 16)[t_index_now_i]
        temporal_resp_plots[[t_index_now_i]] <- 
          plot_heatmap(as_array(temporal_response_1[ , , t_index_now]$cpu()), 
                       do_print = FALSE, do_rasterize = TRUE,
                       use_limits = c(as.numeric(torch_min(temporal_response_1)$cpu()), 
                                      as.numeric(torch_max(temporal_response_1)$cpu()))) + 
          theme(legend.position = "none") +
          ggtitle(paste("t =", round(t_index_now, 2) ))
      }
      temporal_resp_plots <- plot_grid(plotlist = temporal_resp_plots, nrow = 4, align = "hv")
      print(temporal_resp_plots)
      # # test for a single pixel:
      # plot(mat_size_t, as_array(spatial_response_2[80, 100, ]$cpu()), type = "l")
      # lines(mat_size_t, as_array(temporal_response_2[80, 100, ]$cpu()), col = "red")
    } else {
      # save memory:
      rm(resp_fun_gpu, 
         spatial_response_1, spatial_response_2) 
    }
    ## 5. Squaring of zero-phase and half-phase responses
    squared_response <- torch_sqrt(torch_square(temporal_response_1) + torch_square(temporal_response_2))
    if (debug_mode) {
      squared_resp_plots <- vector(mode = "list", length = 16)
      for (t_index_now_i in 1:16) {
        t_index_now <- seq(1, squared_response$size(3), length.out = 16)[t_index_now_i]
        squared_resp_plots[[t_index_now_i]] <- 
          plot_heatmap(as_array(squared_response[ , , t_index_now]$cpu()), 
                       do_print = FALSE, 
                       use_limits = c(0, as.numeric(torch_max(squared_response)$cpu()))) + 
          theme(legend.position = "none") +
          ggtitle(paste("t =", round(t_index_now, 2) ))
      }
      squared_resp_plots <- plot_grid(plotlist = squared_resp_plots, nrow = 4, align = "hv")
      print(squared_resp_plots)
      # max value?
      max(as.vector(as_array(squared_response[ , , t_index_now]$cpu())))
    } else {
      rm(temporal_response_1, temporal_response_2)
    }
    ## 6. Delayed Normalization
    ## The squared response (all positive) will be convoluted with a declining-exponential
    ## low-pass filter. Then, this signal will serve as denominator
    # squared_response is now [y, x, t]
    # get the low-pass filter response function and reshape accordingly
    lp_fun_gpu <- torch_tensor(data = IRF_lp, device = use_device, dtype = preferred_dtype)
    lp_fun_gpu <- lp_fun_gpu$expand(c(squared_response$size(2), 1, lp_fun_gpu$size())) # expand to match matrix dimensions
    lp_fun_gpu <- torch_cat(list(torch_zeros(c(lp_fun_gpu$size(1), lp_fun_gpu$size(2), lp_fun_gpu$size(3)-1), 
                                               device = use_device, dtype = preferred_dtype), # create the zero pad to center the TRF
                                   lp_fun_gpu), dim = 3)
    if (debug_mode) {
      print("Created low-pass temporal response tensor of dimensions for delayed normalization:")
      print(lp_fun_gpu$size())
    }
    # convolve squared responses with low-pass filter
    lowpass_response <- torch_conv1d(input = squared_response, 
                                     weight = lp_fun_gpu$flip(3), 
                                     padding = lp_fun_gpu$size(3)%/%2, 
                                     groups = squared_response$size(2) # is the x dimension
    )
    are_equal(dim(lowpass_response), dim(squared_response))
    # determine the sigma for normalization
    prm.sigma = 0.07 # again, see Figure 11 from "Temporal Dynamics of Neural Responses in Human Visual Cortex"
    prm.n = 1.4
    # perform divisive normalization
    normalized_response <- squared_response
    
    # CONTINUE HERE !!!
    
    # diagnostic plots
    if (debug_mode) {
      dim(normalized_response)
    } else {
      rm(squared_response)
    }
    ## 7. Sum-over-time operation, reduce 3D to 2D [y, x]
    normalized_response_sum <- normalized_response$sum(dim = 3)
    if (debug_mode) {
      print(paste("Computed sum. Resulting dimensions:", paste0(dim(normalized_response_sum), collapse = ",")))
      p_sum <- plot_heatmap(mat = as_array(normalized_response_sum$cpu()), do_print = FALSE)
      print(p_sum)
    }
    ## 8. Pack into final array
    final_2d_over_filters[gabor_list_i, , ] <- as_array(normalized_response_sum$cpu())
    SFs_of_filters[gabor_list_i] <- SF_now
    Oris_of_filters[gabor_list_i] <- Ori_now
    if (output_full_sequences) {
      final_3d_over_filters[gabor_list_i, , , ] <- as_array(normalized_response$cpu())
    }
    # clean up
    rm(normalized_response, normalized_response_sum)
    
  } # end of loop over filters
  # a final matrix of 2D responses?
  if (show_final_maps) {
    all_final_maps <- vector(mode = "list", length = length(gabor_list))
    for (gabor_list_i in 1:length(gabor_list)) {
      all_final_maps[[gabor_list_i]] <- plot_heatmap(mat = final_2d_over_filters[gabor_list_i, , ], 
                                                     do_print = FALSE, do_rasterize = FALSE,
                                                     use_limits = c(0, max(final_2d_over_filters))) + 
        ggtitle(paste0("SF=", round(SFs_of_filters[gabor_list_i], 2), 
                       ", Ori=", round(Oris_of_filters[gabor_list_i], 2) ))
    }
  } else {
    all_final_maps <- NULL
  }
  # return
  return(list(final_2d_over_filters, SFs_of_filters, Oris_of_filters, final_3d_over_filters, all_final_maps))
}
               

```


Try delayed normalization model

```{r}
normSum <- function(x) { x / sum(abs(x)) }

# assume a certain stimulus
t_dn <- seq(0, 1000, by = temporal_resolution)
stim_dn <- rep(0, times = length(t_dn))
stim_dn[t_dn >= 10 & t_dn <= 310] <- 1

# V1 params for Iris Groens model:
prm.tau1 = 0.03 # see Figure 10 of https://doi.org/10.1523/JNEUROSCI.1812-21.2022
prm.weight = 0.4
prm.tau2 = 0.75
prm.sigma = 0.07
prm.n = 1.4
# get Iris Groen's gamma fun (see supplemented code: https://github.com/irisgroen/temporalECoG)
gammaPDF <- function(x, tau, n) {
  y = (x / tau)^(n-1) * exp(-x / tau) / (tau*factorial(n - 1))
  y = y / sum(y)
  return(y)
}
irf_dn_time <- seq(0, 300, by = temporal_resolution)
irf_pos = gammaPDF(irf_dn_time/1000, prm.tau1, 2)
irf_neg = gammaPDF(irf_dn_time/1000, prm.tau1*1.5, 2)
iris_groen_irf = normSum(irf_pos - prm.weight * irf_neg)
plot(irf_dn_time, iris_groen_irf)
# get Iris Groen's declining exponential function - the low-pass filter
iris_groen_irf_norm = normSum(exp(-(irf_dn_time/1000)/prm.tau2))
plot(irf_dn_time, iris_groen_irf_norm)

# 1st convolution + exponentiation
convoluted_iris <- zapsmall(convolve(stim_dn, rev(iris_groen_irf), type = "open"))[1:length(stim_dn)]
convoluted_iris_exp <- abs(convoluted_iris)^prm.n
# 2nd convolution with low-pass filter + exponentiation
convoluted_iris_lowpass <- zapsmall(convolve(convoluted_iris, rev(iris_groen_irf_norm), type = "open"))[1:length(stim_dn)]
convoluted_iris_lowpass_exp <- abs(convoluted_iris_lowpass)^prm.n

# normalization
normalized_iris <- convoluted_iris_exp / ((prm.sigma^prm.n) + convoluted_iris_lowpass_exp)

# plot result
plot(t_dn, stim_dn, type = "l")
lines(t_dn, convoluted_iris_exp, col = "red")
lines(t_dn, convoluted_iris_lowpass_exp, col = "blue")
lines(t_dn, normalized_iris / max(normalized_iris), col = "green")

```







