---
title: "simplified_V1_playground"
author: "Richard Schweitzer"
date: "2023-05-26"
output: github_document
---

Libraries...

```{r}
options(torch.threshold_call_gc = 6000)
library(torch)
library(assertthat)
library(data.table)
library(ggplot2)
library(cowplot)
library(scales)
library(viridis)
library(cowplot)
library(pracma)
library(mgcv)
library(tictoc)
library(minpack.lm)
# deal with massive objects in memory, required: install.packages("BiocManager")
# then: BiocManager::install("DelayedArray")
# library(DelayedArray)
# here: BiocManager::install("HDF5Array")
# library(HDF5Array)
# or here: BiocManager::install("SparseArray")
# library(SparseArray)
```

Screenparameters and more
```{r}
# monitor parameters
monitor.dist <- 180 # distance to monitor in cm
monitor.width <- 200 # width of monitor in cm
monitor.resx <- 960 # horizontal pixels of monitor
monitor.resy <- 540 # vertical pixels of monitor
monitor.screen_center <- data.frame(x = monitor.resx/2, y = monitor.resy/2)
monitor.fps <- 1440
# old and wrong value (monitor.dist was set to 200 instead of 180) for pixels per dva on the screen:
monitor.screen_ppd_old <- 16.7569 
monitor.screen_ppd <- monitor.dist*tan(1*pi/180)/(monitor.width/monitor.resx) # correct
monitor.ppd_correction_factor <- monitor.screen_ppd / monitor.screen_ppd_old
# update stimulus parameters
(sti.freq <- round(c(0.0298/2, 0.0298, 0.0298*2)*monitor.screen_ppd, 2))  # sti.freqp = 0.0298
(sti.sd <- round(8.3778 / monitor.screen_ppd, 2)) # sti.sdp = 8.3778
(sti.sac_amp <- round(275.2756 / monitor.screen_ppd, 2)) # dot.sac_length = 275.2756
c(monitor.resx, monitor.resy) / monitor.screen_ppd
```

Global modeling parameters
```{r}
scr.ppd <- monitor.screen_ppd
spatial_resolution = 1 # resolution must be in pixels, this should be 1, otherwise it's too much computationally.
temporal_resolution = 1000/monitor.fps # temporal resolution of processing function in milliseconds
(SFs = exp(linspace(log(0.25), log(4), 7)))
(Oris = seq(-pi/2, pi/2, length.out = 9)[-1])
Phases = c(0, pi/2)
```

Prepare list of Gabor filters and matching TRFs

```{r}
source("get_IRF_df.R")
irf_space <- get_IRF_df(SFs = SFs, irf_time_range = c(0, 200), 
                        temporal_resolution = temporal_resolution)
# sum for each SF
irf_space[ , .(irf_sum = sum(abs(irf))), by = .(SF)]
# plot the predictions
ggplot(data = irf_space, aes(x = irf_time, y = SF, fill = irf)) + 
  geom_raster(interpolate = TRUE) + 
  coord_cartesian(expand = FALSE) + 
  scale_y_log10() + 
  scale_fill_gradient2() + theme_minimal()
```

... and the matching Gabor filters:
```{r}
source("plot_heatmap.R")
source("get_gabor_field.R")
source("get_gabor_filter_bank.R")
source("get_fft2.R")

# create the Gabor filters
min_SF <- min(SFs)
gabor_list_normal <- get_gabor_filter_bank(
  all_SF = SFs, 
  all_Ori = Oris, # positive means clockwise
  rf_width_override = NaN,# or (1/min_SF) # cpd, set to NaN to let the function determine RF field size
  scr_ppd = scr.ppd, 
  ppd_scaler = 1/spatial_resolution,
  make_gaussian_aperture = TRUE)
# how many have we got?
length(gabor_list_normal)
plot_heatmap(gabor_list_normal[[1]][[1]])
gabor_list_normal[[1]][[3]]
are_equal(round(diff(range(as.numeric(dimnames(gabor_list_normal[[1]][[1]])[[1]]))) / scr.ppd), 
          round(6 * (1/SFs[1]) / 2))


# a sanity check
p_gabor_list_1 <- vector(mode = "list")
p_gabor_list_2 <- vector(mode = "list")
p_gabor_fft <- vector(mode = "list")
for (gi in 1:length(gabor_list_normal)) {
  p_gabor_list_1[[gi]] <- plot_heatmap(mat = gabor_list_normal[[gi]][[1]], do_print = FALSE) + 
    theme(legend.position = "none") + theme_nothing()
  p_gabor_list_2[[gi]] <- plot_heatmap(mat = gabor_list_normal[[gi]][[2]], do_print = FALSE) + 
    theme(legend.position = "none") + theme_nothing()
  # fft:
  gabor_fft <- get_fft2(stim_image_gray = gabor_list_normal[[gi]][[1]], scr.ppd = scr.ppd, pad_image_size = 200)
  p_gabor_fft[[gi]] <- plot_heatmap(mat = gabor_fft[[4]], do_print = FALSE)  + 
    theme(legend.position = "none") + theme_nothing()
  rm(gabor_fft)
}
# plot 1: zero phase
plot_grid(plotlist = p_gabor_list_1, nrow = length(SFs))
rm(p_gabor_list_1)
# plot 2: half phase
plot_grid(plotlist = p_gabor_list_2, nrow = length(SFs))
rm(p_gabor_list_2)
# plot 3: fft
plot_grid(plotlist = p_gabor_fft, nrow = length(SFs))
rm(p_gabor_fft)
```

Now Heiko Schuett's Gabor function:

```{r}
source("get_log_gabor_heiko.R")
source("get_gabor_field_heiko.R")

gabor_list_heiko <- get_gabor_filter_bank(
  use_log_gabor_heiko = TRUE,
  all_SF = SFs, 
  all_Ori = Oris, # positive means clockwise
  rf_width_override = (1/SFs) / 2 * 8, # let's say eight times the Gaussian SD
  scr_ppd = scr.ppd, 
  ppd_scaler = 1/spatial_resolution,
  make_gaussian_aperture = TRUE)
length(gabor_list_heiko)

## visual comparison of Gabors
gabor_comparison_nr <- 4
# Heiko's log-Gabor
gabor_heiko <- gabor_list_heiko[[gabor_comparison_nr]][[1]]
dimnames(gabor_heiko) <- NULL
# the normal Gabor
gabor_normal <- gabor_list_normal[[gabor_comparison_nr]][[1]]
gabor_normal_adj <- matrix(0, nrow = nrow(gabor_heiko), ncol = ncol(gabor_heiko))
gabor_normal_adj[round(nrow(gabor_normal_adj)/2-nrow(gabor_normal)/2):round(nrow(gabor_normal_adj)/2-
                                                                              nrow(gabor_normal)/2+nrow(gabor_normal)-1), 
                 round(ncol(gabor_normal_adj)/2-ncol(gabor_normal)/2):round(ncol(gabor_normal_adj)/2-
                                                                              ncol(gabor_normal)/2+ncol(gabor_normal)-1)] <- 
  gabor_normal
dimnames(gabor_normal_adj) <- NULL
# plot them together
plot_grid(plot_heatmap(gabor_heiko), 
          plot_heatmap(gabor_normal_adj), 
          ncol = 2)
plot_grid(plot_heatmap(abs(get_fft2(gabor_heiko, scr.ppd = scr.ppd)[[4]])) + 
            geom_vline(xintercept = gabor_list_heiko[[gabor_comparison_nr]][[3]]['rf_freq_dva'], linetype = "dotted", linewidth = 0.4), 
          plot_heatmap(abs(get_fft2(gabor_normal_adj, scr.ppd = scr.ppd)[[4]])) + 
            geom_vline(xintercept = gabor_list_heiko[[gabor_comparison_nr]][[3]]['rf_freq_dva'], linetype = "dotted", linewidth = 0.4), 
          ncol = 2)

# Gabor matrix sizes - comparison
rbind(unlist(lapply(X = gabor_list_heiko, FUN = function(x) { nrow(x[[1]]) })), 
      unlist(lapply(X = gabor_list_normal, FUN = function(x) { nrow(x[[1]]) }))
)

# a sanity check
p_gabor_list_1 <- vector(mode = "list")
p_gabor_list_2 <- vector(mode = "list")
p_gabor_fft <- vector(mode = "list")
for (gi in 1:length(gabor_list_heiko)) {
  p_gabor_list_1[[gi]] <- plot_heatmap(mat = gabor_list_heiko[[gi]][[1]], do_print = FALSE) + 
    theme(legend.position = "none") + theme_nothing()
  p_gabor_list_2[[gi]] <- plot_heatmap(mat = gabor_list_heiko[[gi]][[2]], do_print = FALSE) + 
    theme(legend.position = "none") + theme_nothing()
  # fft:
  gabor_fft <- get_fft2(stim_image_gray = gabor_list_heiko[[gi]][[1]], 
                        scr.ppd = scr.ppd, pad_image_size = 512)
  p_gabor_fft[[gi]] <- plot_heatmap(mat = gabor_fft[[4]], do_print = FALSE)  + 
    theme(legend.position = "none") + theme_nothing()
  rm(gabor_fft)
}
# plot 1: zero phase
plot_grid(plotlist = p_gabor_list_1, nrow = length(SFs))
rm(p_gabor_list_1)
# plot 2: half phase
plot_grid(plotlist = p_gabor_list_2, nrow = length(SFs))
rm(p_gabor_list_2)
# plot 3: fft
plot_grid(plotlist = p_gabor_fft, nrow = length(SFs))
rm(p_gabor_fft)
```


Now try delayed normalization model

```{r}
source("normSum.R")

# assume a certain stimulus
t_dn <- seq(0, 1000, by = temporal_resolution)
stim_dn <- rep(0, times = length(t_dn))
stim_dn[t_dn >= 10 & t_dn <= 310] <- 1

# V1 params for Iris Groens model:
prm.tau1 = 0.03 # see Figure 10 of https://doi.org/10.1523/JNEUROSCI.1812-21.2022
prm.weight = 0.4
prm.tau2 = 0.75
prm.sigma = 0.07
prm.n = 1.4
# get Iris Groen's gamma fun (see supplemented code: https://github.com/irisgroen/temporalECoG)
gammaPDF <- function(x, tau, n) {
  y = (x / tau)^(n-1) * exp(-x / tau) / (tau*factorial(n - 1))
  y = y / sum(y)
  return(y)
}
irf_dn_time <- seq(0, 300, by = temporal_resolution)
irf_pos = gammaPDF(irf_dn_time/1000, prm.tau1, 2)
irf_neg = gammaPDF(irf_dn_time/1000, prm.tau1*1.5, 2)
iris_groen_irf = normSum(irf_pos - prm.weight * irf_neg)
plot(irf_dn_time, iris_groen_irf)
# get Iris Groen's declining exponential function - the low-pass filter
iris_groen_irf_norm = normSum(exp(-(irf_dn_time/1000)/prm.tau2))
plot(irf_dn_time, iris_groen_irf_norm)

# 1st convolution + exponentiation
convoluted_iris <- zapsmall(convolve(stim_dn, rev(iris_groen_irf), type = "open"))[1:length(stim_dn)]
convoluted_iris_exp <- abs(convoluted_iris)^prm.n
# 2nd convolution with low-pass filter + exponentiation
convoluted_iris_lowpass <- zapsmall(convolve(convoluted_iris, rev(iris_groen_irf_norm), type = "open"))[1:length(stim_dn)]
convoluted_iris_lowpass_exp <- abs(convoluted_iris_lowpass)^prm.n

# normalization
normalized_iris <- convoluted_iris_exp / ((prm.sigma^prm.n) + convoluted_iris_lowpass_exp)

# plot result
plot(t_dn, stim_dn, type = "l")
lines(t_dn, convoluted_iris_exp, col = "red")
lines(t_dn, convoluted_iris_lowpass_exp, col = "blue")
lines(t_dn, normalized_iris / max(normalized_iris), col = "green")

```


# Test the model 

Now load a random stimulus from SDEC and create arbitrary trajectory
```{r}
source("load_noise_patch.R")
all_noise_patch_paths <- dir("stimuli_matrices", pattern = "*.csv")
# load them all
all_noise_patches <- vector(mode = "list", length = length(all_noise_patch_paths))
all_noise_patch_plots <- vector(mode = "list", length = length(all_noise_patch_paths))
for (noise_patch_path_i in 1:length(all_noise_patch_paths)) {
  noise_patch_path <- all_noise_patch_paths[noise_patch_path_i]
  all_noise_patches[[noise_patch_path_i]] <- load_noise_patch(stim_name = noise_patch_path)
  all_noise_patch_plots[[noise_patch_path_i]] <- plot_heatmap(mat = all_noise_patches[[noise_patch_path_i]]) + 
    ggtitle(noise_patch_path)
}
plot_grid(plotlist = all_noise_patch_plots, align = "hv")
# what's the size of the noise patch?
dim(all_noise_patches[[1]]) / scr.ppd
sti.sd * 6 # this is their specified size
```

Get all saccade data and fit by model. 
```{r}
# do the preprocessing of saccades again? if so, set to TRUE:
do_overwrite_all_saccades <- FALSE

if (!file.exists("all_saccades.rda") || do_overwrite_all_saccades) { # process all saccades again? (slow)
  
  source("five_point_smoother.R")
  source("resample_signal_df.R")
  
  # this is the path to the cleaned eye+EEG data
  path_to_saccade_data <- "/home/richard/Seafile/Intrasaccadic VEPs/SDEC/SDEC DATA/Data"
  load(file = file.path(path_to_saccade_data, "SDEC_data_cleaned_rej2.rda"))
  rm(SDEC_eeg, SDEC_eeg_event, SDEC_eeg_timings, SDEC_data, sdec2)
  
  # go through all trials and extract primary saccade
  setkeyv(SDEC_sac_raw, "ID")
  setkeyv(sdec, "ID")
  all_IDs <- unique(sdec$ID) 
  are_equal(sort(all_IDs), sort(unique(SDEC_sac_raw$ID))) # make sure everything is complete
  do_fits <- FALSE
  assert_that(do_fits == FALSE) # code not prepared for fits (anymore)
  if (do_fits) {
    source("compressed_exp.R")
    source("harmonic_osc.R")
    saccade_params <- vector(mode = "list", length = length(all_IDs))
  } else {
    all_saccades <- vector(mode = "list", length = length(all_IDs))
  }
  # go through all trials
  for (id_i in 1:length(all_IDs)) {
    # get data 
    id <- all_IDs[id_i]
    trial <- sdec[.(id)]
    assert_that(nrow(trial)==1)
    sac <- SDEC_sac_raw[.(id)]
    assert_that(nrow(sac)>100)
    sac[ , time_sac_on := time - trial$primary_onset]
    sac[ , time_sac_off := time - trial$primary_offset]
    sac[ , time_stim_on := time - trial$stim_on_time]
    sac[ , time_secondary_on := time - trial$secondary_onset]
    # extract and add factors
    if (any(is.na(sac$time_secondary_on))) {
      sac <- sac[time_sac_on>=(-100) & time_sac_on<=120]
    } else {
      sac <- sac[time_sac_on>=(-100) & time_secondary_on<=(-5)]
    }
    sac[ , start_left_f := trial$start_left_f]
    sac[ , streak_present_f := trial$streak_present_f]
    sac[ , move_direction_f := trial$move_direction_f]
    # apply smoothing
    sac[ , X := five_point_smoother(X)]
    sac[ , Y := five_point_smoother(Y)]
    # upsample
    #  plot(sac$time_sac_on, sac$X)
    sac <- resample_signal_df(df = sac, 
                              time_col = "time", 
                              id_cols = c("ID", "subj_id", "session_id", "start_left_f", "streak_present_f", "move_direction_f"), 
                              constant_cols = c(),
                              signal_cols = c("X", "Y", "PUPIL"), 
                              linear_cols = c("time_cue_on", "time_stim_on", "time_sac_on", "time_sac_off"), 
                              new_samp_rate = monitor.fps, 
                              plot_results = FALSE, debug_mode = FALSE)
    #  points(sac$time_sac_on, sac$X, col = "red")
    assert_that(round(median(diff(sac$time), na.rm = TRUE), 3)==round(temporal_resolution, 3))
    # for replay: get stimulus position over time
    stim_on_here <- which.min(abs(sac$time_stim_on))
    stim_frames <- round(trial$stim_dur*1000/temporal_resolution)
    assert_that(stim_frames==36) # this is always the case
    setDT(sac)
    # static stimulus:
    sac[ , static_stim_x := trial$stim_pos_x_start]
    sac[ , static_stim_y := trial$stim_pos_y_start]
    # moving stimulus:
    sac[1:stim_on_here, stim_x := trial$stim_pos_x_start]
    sac[1:stim_on_here, stim_y := trial$stim_pos_y_start]
    sac[ , stim_moving := FALSE]
    sac[stim_on_here:(stim_on_here+stim_frames-1), 
        stim_moving := TRUE]
    sac[stim_on_here:(stim_on_here+stim_frames-1), 
        stim_x := seq(trial$stim_pos_x_start, trial$stim_pos_x_final, 
                      length.out = stim_frames)]
    sac[stim_on_here:(stim_on_here+stim_frames-1), 
        stim_y := seq(trial$stim_pos_y_start, trial$stim_pos_y_final, 
                      length.out = stim_frames)]
    sac[(stim_on_here+stim_frames):nrow(sac), stim_x := trial$stim_pos_x_final]
    sac[(stim_on_here+stim_frames):nrow(sac), stim_y := trial$stim_pos_y_final]
    # compute retinal coordinates:
    sac[ , retinal_x := stim_x - X]
    sac[ , retinal_y := stim_y - Y]
    sac[ , retinal_x_static := static_stim_x - X]
    sac[ , retinal_y_static := static_stim_y - Y]
    # fit harmonic oscillator to the distance to catch PSO ???? - not needed
    if (!do_fits) {
      # add saccade to our list
      all_saccades[[id_i]] <- sac
    } else {
      sac$dist <- sqrt((sac$X-sac$X[1])^2 + (sac$Y-sac$Y[1])^2 ) / scr.ppd
      sac$dir <- atan2(sac$Y-sac$Y[1], sac$X-sac$X[1]) #### * 180/pi
      mean_sac_dir <- mean(sac$dir[sac$time_sac_off>=0])
      # rotate so that we capture the orthogonal deviation
      sac$x_r <- (sac$X-sac$X[1]) * cos(-mean_sac_dir) - (sac$Y-sac$Y[1]) * sin(-mean_sac_dir) + sac$X[1]
      sac$y_r <- (sac$X-sac$X[1]) * sin(-mean_sac_dir) + (sac$Y-sac$Y[1]) * cos(-mean_sac_dir) + sac$Y[1]
      sac$dir_control <- atan2(sac$y_r-sac$y_r[1], sac$x-sac$x_r[1]) 
      # go fit?
      fit_worked <- FALSE
      try({
        hof <- nlsLM(data = sac, formula = dist ~ harmonic_osc(time_sac_on, delay, amp, dur, omega), 
                     start = c(delay = 0, amp = max(sac$dist), 
                               dur = sac$time_sac_on[sac$time_sac_off>=0][1]/2, 
                               omega = 0.05), 
                     lower = c(delay = min(sac$time_sac_on), amp = 0, dur = 0, omega = 0),
                     control = nls.lm.control(maxiter = 200))
        coef_hof <- data.table(t(coef(hof)))
        coef_hof[ , subj_id := trial$subj_id]
        coef_hof[ , ID := trial$ID]
        coef_hof[ , start_left := trial$start_left_f]
        coef_hof[ , streak_present := trial$streak_present_f]
        coef_hof[ , move_direction := trial$move_direction_f]
        # fit polynomial to saccade curvature
        curv <- lm(data = sac, formula = y_r ~ poly(time_sac_on,3)) # cubic
        coef_curv <- data.table(t(coef(curv)))
        colnames(coef_curv) <- c("d", "c", "b", "a")
        # combine the two
        params_now <- cbind(coef_curv, coef_hof)
        saccade_params[[id_i]] <- params_now
        fit_worked <- TRUE
        # # plot PSO fit?
        # plot(sac$time_sac_on, sac$dist)
        # lines(sac$time_sac_on, predict(hof))
        # # plot curvature fit?
        # plot(sac$time_sac_on, sac$y_r)
        # lines(sac$time_sac_on, predict(curv))
      }, silent = TRUE)
      #if (!fit_worked) { print(paste0("Could not fit at id_i=", id_i)) }
    } 
  }
  # make data table
  if (do_fits) {
    saccade_params <- rbindlist(saccade_params)
  } else {
    all_saccades <- rbindlist(all_saccades)
    all_saccades <- all_saccades[!is.na(time)]
  }
  # clean up the saccade data
  rm(SDEC_sac_raw)
  
  # plot all retinal trajectories?
  for (subj_now in unique(all_saccades$subj_id)) {
    p_retinal_subj <- ggplot(data = all_saccades[subj_id==subj_now], 
                             aes(x = retinal_x, y = retinal_y, 
                                 color = stim_moving, group = paste(ID, stim_moving) )) + 
      geom_point(size = 0.1, alpha = 0.1) + 
      facet_wrap(~start_left_f+move_direction_f, nrow = 2, scales = "free") + 
      theme_minimal() + ggtitle(subj_now)
    print(p_retinal_subj)
  }
  rm(p_retinal_subj)
  
  # save this:
  save(list = c("all_saccades", "sdec"), file = "all_saccades.rda", compress = "xz")
  
} else { # the file is there:
  
  # load all the saccades and their retinal trajectories
  load("all_saccades.rda")
  
}

# pick a random saccade
random_sac_ID <- sample(unique(all_saccades[move_direction_f=="inward" & start_left_f=="rightward saccade", ID]))[1]
random_sac <- all_saccades[ID==random_sac_ID]
x_stim_traj <- random_sac$retinal_x
y_stim_traj <- random_sac$retinal_y
x_static_stim_traj <- random_sac$retinal_x_static
y_static_stim_traj <- random_sac$retinal_y_static
t_stim_traj <- random_sac$time_cue_on-random_sac$time_cue_on[1]
plot(x_stim_traj, y_stim_traj)
points(x_static_stim_traj, y_static_stim_traj, col = "red")
plot(t_stim_traj, sqrt((x_stim_traj-x_stim_traj[1])^2 + (y_stim_traj-y_stim_traj[1])^2))
points(t_stim_traj, sqrt((x_static_stim_traj-x_static_stim_traj[1])^2 + (y_static_stim_traj-y_static_stim_traj[1])^2), col = "red")
```

This is the model: 

```{r}
# to debug:
# signal_x = x_stim_traj[200:250]
# signal_y = y_stim_traj[200:250]
# signal_t = t_stim_traj[200:250]
# gabor_list = gabor_list_heiko
# irf_df = irf_space
# stim_mat = all_noise_patches[[1]]


# source the two spatial convolution functions here:
source("standard_conv_with_torch_3d.R")
source("fft_conv_with_torch_3d.R")
source("plot_heatmap.R")
source("temporal_conv_kelly_fft.R")
source("kelly_vel.R")

# this is the V1 module
source("v1.R")
```

Test this function, let it plot, etc.
```{r}
# normal Gabors with normalization pool
v1_output <- v1(stim_mat = all_noise_patches[[1]], # the noise patch as a matrix
                gabor_list = gabor_list_normal, # a list of gabor filters created by 'get_gabor_filter_bank'
                irf_df = irf_space, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                signal_x = x_stim_traj, 
                signal_y = y_stim_traj, 
                signal_t = t_stim_traj, # properties of the signal (relative to saccade onset)
                output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                use_half_precision = TRUE, # if a GPU is available, use half-precision?
                spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                use_normalization_pool = TRUE, # use normalization pool?
                debug_mode = FALSE, # shows output at every step
                show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                final_maps_filename = "v1_output_normal_gabor_normalizepool.pdf" # name of the pdf file if show_final_maps==TRUE
)
# normal Gabors WITHOUT normalization pool
v1_output <- v1(stim_mat = all_noise_patches[[1]], # the noise patch as a matrix
                gabor_list = gabor_list_normal, # a list of gabor filters created by 'get_gabor_filter_bank'
                irf_df = irf_space, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                signal_x = x_stim_traj, 
                signal_y = y_stim_traj, 
                signal_t = t_stim_traj, # properties of the signal (relative to saccade onset)
                output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                use_half_precision = TRUE, # if a GPU is available, use half-precision?
                spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                use_normalization_pool = FALSE, # use normalization pool?
                debug_mode = FALSE, # shows output at every step
                show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                final_maps_filename = "v1_output_normal_gabor.pdf" # name of the pdf file if show_final_maps==TRUE
)
# Heiko's log Gabors with normalization pool
v1_output <- v1(stim_mat = all_noise_patches[[1]], # the noise patch as a matrix
                gabor_list = gabor_list_heiko, # a list of gabor filters created by 'get_gabor_filter_bank'
                irf_df = irf_space, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                signal_x = x_stim_traj, 
                signal_y = y_stim_traj, 
                signal_t = t_stim_traj, # properties of the signal (relative to saccade onset)
                output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                use_half_precision = TRUE, # if a GPU is available, use half-precision?
                spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                use_normalization_pool = TRUE, # use normalization pool?
                debug_mode = FALSE, # shows output at every step
                show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                final_maps_filename = "v1_output_log_gabor_normalizepool.pdf" # name of the pdf file if show_final_maps==TRUE
)
# Heiko's log Gabors WITHOUT normalization pool
v1_output <- v1(stim_mat = all_noise_patches[[1]], # the noise patch as a matrix
                gabor_list = gabor_list_heiko, # a list of gabor filters created by 'get_gabor_filter_bank'
                irf_df = irf_space, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                signal_x = x_stim_traj, 
                signal_y = y_stim_traj, 
                signal_t = t_stim_traj, # properties of the signal (relative to saccade onset)
                output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                use_half_precision = TRUE, # if a GPU is available, use half-precision?
                spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                use_normalization_pool = FALSE, # use normalization pool?
                debug_mode = FALSE, # shows output at every step
                show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                final_maps_filename = "v1_output_log_gabor.pdf" # name of the pdf file if show_final_maps==TRUE
)
```

# Let the model run on a controlled comparison of all move direction vs static

```{r}
# # test whether one could index a RleArray directly
# require(DelayedArray)
# required_dims <- c(10, 20, 30, 1000)
# dat <- Rle(values = NaN, lengths = prod(required_dims))
# A <- RleArray(dat, dim=required_dims)
# dim(A)
# print(object.size(A))
# # iteratively assign
# for (it in 1:required_dims[1]) {
#   A[it, , , ] <- array(data = rnorm(n = prod(required_dims[2:4])), dim = required_dims[2:4])
#   print(object.size(A))
# }
# A[1, 1:10, 1:10, 1]
# A <- as(A, "RleArray")
# A[1, 1:10, 1:10, 1]
# print(object.size(A))


source("compute_difference_3d_matrices.R")
n_trials_per_cell <- 1

# run through all conditions
all_2d_model_predictions <- NULL
all_3d_model_predictions <- NULL
for (start_left_now in unique(all_saccades$start_left_f)) {
  for (subj_now in unique(all_saccades$subj_id)[1]) {
    # get the corresponding target noise patch
    noise_patch_now <- all_noise_patches[[as.numeric(subj_now)]]
    # get a set of random trials in that condition
    sampled_IDs <- sample(unique(all_saccades[move_direction_f!="static" & start_left_f==start_left_now & subj_id==subj_now, 
                                              ID]))[1:n_trials_per_cell]
    # run through those
    for (n_trial_now in 1:n_trials_per_cell) {
      cond_def <- paste(start_left_now, subj_now, n_trial_now, sep = ", ")
      print(cond_def)
      # get trial data
      ID_now <- sampled_IDs[n_trial_now]
      trial_sac <- all_saccades[ID==ID_now]
      trial_data <- sdec[ID==ID_now]
      
      # iterate over possible movement directions
      for (move_direction_now in c("static", as.character(unique(all_saccades[move_direction_f!="static", move_direction_f]))) ) {
        # extend of stimulus movement
        stim_travel_dist <- sqrt((tail(trial_sac[stim_moving==TRUE, stim_x], 1)-
                                    head(trial_sac[stim_moving==TRUE, stim_x], 1))^2 + 
                                   (tail(trial_sac[stim_moving==TRUE, stim_y], 1)-
                                      head(trial_sac[stim_moving==TRUE, stim_y], 1))^2)
        stim_frames <- sum(trial_sac$stim_moving)
        # in case we have a moving stimulus...
        if (move_direction_now != "static") {
          staticmoving_cond <- "moving"
          # now, regardless of the true movement direction in the trial, reconstruct retinal trajectories 
          stim_on_here <- min(which(trial_sac$stim_moving==TRUE))
          stim_pos_x_start <- trial_sac[stim_on_here, stim_x]
          stim_pos_y_start <- trial_sac[stim_on_here, stim_y]
          # decide where the stimulus should go
          if ((move_direction_now=="inward" & start_left_now=="rightward saccade") | 
              (move_direction_now=="outward" & start_left_now=="leftward saccade")) {
            stim_pos_x_final <- stim_pos_x_start - stim_travel_dist
            stim_pos_y_final <- stim_pos_y_start
          } else if ((move_direction_now=="inward" & start_left_now=="leftward saccade") | 
                     (move_direction_now=="outward" & start_left_now=="rightward saccade")) {
            stim_pos_x_final <- stim_pos_x_start + stim_travel_dist
            stim_pos_y_final <- stim_pos_y_start
          } else if (move_direction_now=="upward") {
            stim_pos_x_final <- stim_pos_x_start
            stim_pos_y_final <- stim_pos_y_start - stim_travel_dist
          } else if (move_direction_now=="downward") {
            stim_pos_x_final <- stim_pos_x_start
            stim_pos_y_final <- stim_pos_y_start + stim_travel_dist
          } else {
            stop("Check move_direction_now!")
          }
          # overwrite the trial's stimulus trajectory
          trial_sac[stim_on_here:(stim_on_here+stim_frames-1), 
                    stim_x := seq(stim_pos_x_start, stim_pos_x_final, 
                                  length.out = stim_frames)]
          trial_sac[stim_on_here:(stim_on_here+stim_frames-1), 
                    stim_y := seq(stim_pos_y_start, stim_pos_y_final, 
                                  length.out = stim_frames)]
          # ... and its endpoint
          trial_sac[(stim_on_here+stim_frames):nrow(trial_sac), stim_x := stim_pos_x_final]
          trial_sac[(stim_on_here+stim_frames):nrow(trial_sac), stim_y := stim_pos_y_final]
          # and accordingly compute retinal coordinates:
          trial_sac[ , retinal_x := stim_x - X]
          trial_sac[ , retinal_y := stim_y - Y]
        } else { # in case we have a static stimulus
          staticmoving_cond <- "static"
          trial_sac[ , retinal_x := retinal_x_static ]
          trial_sac[ , retinal_y := retinal_y_static ]
        }
        # remove samples were target is moving to simulate the absent condition
        trial_sac_absent <- copy(trial_sac)
        trial_sac_absent <- trial_sac_absent[stim_moving==FALSE]
        # check?
        par(mfrow=c(1,2))
        plot(trial_sac$time_stim_on, trial_sac$retinal_x, 
             main = paste(cond_def, move_direction_now, sep = ", "))
        points(trial_sac_absent$time_stim_on, trial_sac_absent$retinal_x, col = "red")
        plot(trial_sac$retinal_x, trial_sac$retinal_y, 
             main = paste(cond_def, move_direction_now, sep = ", "))
        points(trial_sac_absent$retinal_x, trial_sac_absent$retinal_y, col = "red")
        par(mfrow=c(1,1))
        
        # run model twice, once for present, once for absent conditions
        final_2d_over_filters <- vector(mode = "list", length = 2)
        final_3d_over_filters <- vector(mode = "list", length = 2)
        for (model_cond_i in 1:2) { # iterate over present and absent
          # in what condition?
          if (model_cond_i==1) { 
            model_cond <- "present"
            trial_sac_test <- trial_sac
          } else {
            model_cond <- "absent"
            trial_sac_test <- trial_sac_absent
          }
          # run model
          print(paste(move_direction_now, model_cond))
          v1_output <- v1(stim_mat = noise_patch_now, # the noise patch as a matrix
                          gabor_list = gabor_list_heiko, # a list of gabor filters created by 'get_gabor_filter_bank'
                          normalize_RFs = TRUE, 
                          irf_df = irf_space, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                          normalize_IRFs = TRUE, 
                          signal_x = trial_sac_test$retinal_x, 
                          signal_y = trial_sac_test$retinal_y, 
                          signal_t = trial_sac_test$time_sac_off, 
                          output_full_sequences = TRUE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                          output_full_sequences_use_RleArray = FALSE, 
                          signal_x_range = round(c(min(trial_sac_test$retinal_x), 
                                                   max(trial_sac_test$retinal_x))), 
                          signal_y_range = round(c(min(trial_sac_test$retinal_y), 
                                                   max(trial_sac_test$retinal_y))), 
                          signal_t_range = c(min(trial_sac_test$time_sac_off), 
                                             max(trial_sac_test$time_sac_off)), # the min and max of temporal scale
                          no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                          use_half_precision = TRUE, # if a GPU is available, use half-precision?
                          spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                          temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                          normalize_override = FALSE, 
                          use_normalization_pool = TRUE, # use normalization pool?
                          debug_mode = FALSE, # shows output at every step
                          show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                          final_maps_filename = paste0("v1_output_current_", model_cond, 
                                                       "_", staticmoving_cond, ".pdf") # name of the pdf file if show_final_maps==TRUE
          )
          # get Oris and SFs
          SFs_of_filters <- v1_output[[2]]
          Oris_of_filters <- v1_output[[3]]
          
          if (!is.null(v1_output[[4]])) {
            do_3d_stuff <- TRUE
          } else {
            do_3d_stuff <- FALSE
          }
          ## 1. summed-over-time model responses
          # create data table from summed output
          df_final_2d <- as.data.table(v1_output[[1]])
          colnames(df_final_2d) <- c("filter_id", "y", "x", 
                                     paste("sum_resp", model_cond, sep = "_"))
          SF_col <- vector(mode = "numeric", length = nrow(df_final_2d))
          Ori_col <- vector(mode = "numeric", length = nrow(df_final_2d))
          for (r in 1:nrow(df_final_2d)) {
            SF_col[r] <- SFs_of_filters[df_final_2d$filter_id[r]]
            Ori_col[r] <- Oris_of_filters[df_final_2d$filter_id[r]]
          }
          df_final_2d[ , SF := SF_col]
          df_final_2d[ , Ori := Ori_col]
          
          ## 2.1 reduce size of 3d matrix, by using compression
          if (do_3d_stuff) {
            use_delayed_array <- "RleArray" # "HDF5Array" or "RleArray"
            if (use_delayed_array == "SparseArray") { package_name <- "SparseArray" }
            if (use_delayed_array == "RleArray") { package_name <- "DelayedArray" }
            if (use_delayed_array == "HDF5Array") { package_name <- "HDF5Array" }
            require(package_name, character.only = TRUE)
            print(paste("Compressing to", use_delayed_array, "..."))
            if (staticmoving_cond=="static") {
              if (model_cond=="present") {
                if (is(v1_output[[4]], use_delayed_array)) {
                  v1_output_3d_static_present <- v1_output[[4]]
                } else {
                  v1_output_3d_static_present <- as(v1_output[[4]], use_delayed_array) 
                }
              } else {
                if (is(v1_output[[4]], use_delayed_array)) {
                  v1_output_3d_static_absent <- v1_output[[4]]
                } else {
                  v1_output_3d_static_absent <- as(v1_output[[4]], use_delayed_array) 
                }
              }
            } else if (staticmoving_cond=="moving") {
              if (model_cond=="present") {
                if (is(v1_output[[4]], use_delayed_array)) {
                  v1_output_3d_moving_present <- v1_output[[4]]
                } else {
                  v1_output_3d_moving_present <- as(v1_output[[4]], use_delayed_array) 
                }
              } else {
                if (is(v1_output[[4]], use_delayed_array)) {
                  v1_output_3d_moving_absent <- v1_output[[4]]
                } else {
                  v1_output_3d_moving_absent <- as(v1_output[[4]], use_delayed_array)
                }
              }
            }
            print("Done.")
            # remove that massive list from memory
            rm(v1_output)
            gc() # ... and clean up
          }
          
          ## 2.2 model responses resolved over time 
          df_diff_resp_over_time <- NULL
          if (do_3d_stuff & staticmoving_cond=="moving") {
            # select the right arrays
            if (model_cond=="present") {
              v1_output_3d_moving <- v1_output_3d_moving_present 
              v1_output_3d_static <- v1_output_3d_static_present
              # # to debug:
              # mat_1 <- v1_output_3d_moving[1, , , ]
              # mat_2 <- v1_output_3d_static[1, , , ]
            } else {
              v1_output_3d_moving <- v1_output_3d_moving_absent
              v1_output_3d_static <- v1_output_3d_static_absent
            }
            print("Computing difference between moving and static from delayed arrays...")
            # Parallelize here. Inspired by: https://www.glennklockwood.com/data-intensive/r/foreach-parallelism.html
            require(foreach)
            if (use_delayed_array!="SparseArray") {
              n_cores <- length(unique(SFs_of_filters)) # use parallelization not for sparse arrays - too memory intense
              require(doMC)
              registerDoMC(length(unique(SFs_of_filters)))
              `%dofun%` <- `%dopar%`
            } else {
              `%dofun%` <- `%do%`
            }
            # compute difference by running through filters (means and sums take a long time with RleArrays)
            df_diff_resp_over_time <- 
              foreach(filter_i = 1:dim(v1_output_3d_static)[1], .verbose = FALSE, .errorhandling = "stop", 
                      .combine = 'rbind', 
                      .packages = c("assertthat", "data.table", package_name) ) %dofun% {
                        print(paste("Computing means ... ", filter_i, "of", dim(v1_output_3d_static)[1]))
                        # current SF and Orientation
                        SF_now <- SFs_of_filters[filter_i]
                        Ori_now <- Oris_of_filters[filter_i]
                        ## from delayed/sparse array get:
                        # compute differences between the two 3d matrices
                        v1_output_3d_filter_diff_sums_over_time <- 
                          compute_difference_3d_matrices(mat_1 = v1_output_3d_moving[filter_i, , , ], 
                                                         mat_2 = v1_output_3d_static[filter_i, , , ], 
                                                         use_RLE = TRUE, 
                                                         absolute_difference = TRUE)
                        # absolute means
                        v1_output_3d_filter_moving_sums_over_time <- apply(X = v1_output_3d_moving[filter_i, , , ], 
                                                                           MARGIN = 3,
                                                                           FUN = mean)
                        v1_output_3d_filter_static_sums_over_time <- apply(X = v1_output_3d_static[filter_i, , , ], 
                                                                           MARGIN = 3,
                                                                           FUN = mean)
                        are_equal(names(v1_output_3d_filter_moving_sums_over_time), names(v1_output_3d_filter_diff_sums_over_time))
                        are_equal(names(v1_output_3d_filter_moving_sums_over_time), names(v1_output_3d_filter_static_sums_over_time))
                        # combine
                        df_now <- data.table(SF = SF_now, Ori = Ori_now, 
                                             time = as.numeric(names(v1_output_3d_filter_diff_sums_over_time)), 
                                             diff_resp_sum = as.numeric(v1_output_3d_filter_diff_sums_over_time), 
                                             moving_resp_sum = as.numeric(v1_output_3d_filter_moving_sums_over_time), 
                                             static_resp_sum = as.numeric(v1_output_3d_filter_static_sums_over_time) )
                        colnames(df_now)[grepl(x = colnames(df_now), pattern = "_resp_sum")] <- 
                          paste(colnames(df_now)[grepl(x = colnames(df_now), pattern = "_resp_sum")], 
                                model_cond, sep = "_")
                        # save
                        return(df_now)
                      } # end of iterate over filter_i
            # clean up
            rm(v1_output_3d_moving, v1_output_3d_static)
            rm(SF_col, Ori_col, SFs_of_filters, Oris_of_filters)
            gc()
            print("Done.")
            
          } # end of if (do_3d_stuff & staticmoving_cond=="moving")
          
          # save this
          final_2d_over_filters[[model_cond_i]] <- df_final_2d
          final_3d_over_filters[[model_cond_i]] <- df_diff_resp_over_time
          
        } # now we have model outputs to both present and absent cases
        
        #asdf # this stops
        
        # 3D results: merge absent and present results and save them
        if (staticmoving_cond=="moving") {
          are_equal(sort(unique(final_3d_over_filters[[1]]$time)), sort(unique(final_3d_over_filters[[2]]$time)))
          final_3d_over_filters <- merge.data.table(x = final_3d_over_filters[[1]], 
                                                    y = final_3d_over_filters[[2]], 
                                                    by = c("SF", "Ori", "time"))
          final_3d_over_filters[ , move_direction_f := move_direction_now]
          final_3d_over_filters[ , subj_id := subj_now]
          final_3d_over_filters[ , start_left_f := start_left_now]
          final_3d_over_filters[ , ID := ID_now]
          
          # TO DO: plot output to present-static vs absent-static
          p_diff_3d_over_filters <- ggplot(data = final_3d_over_filters, 
                                           aes(x = time, y = diff_resp_sum_absent)) + 
            geom_line(aes(color = "1 absent")) + 
            geom_line(data = final_3d_over_filters, 
                      aes(x = time, y = diff_resp_sum_present, color = "2 present")) + 
            geom_line(data = final_3d_over_filters, 
                      aes(x = time, y = -1*(diff_resp_sum_present-diff_resp_sum_absent), color = "3 difference")) + 
            facet_grid(round(SF, 2)~round(Ori, 2)) + 
            theme_minimal() + 
            ggtitle(paste(cond_def, sep = ", ")) + 
            scale_color_viridis_d(end = 1)
          print(p_diff_3d_over_filters)
          
          # save 3D data
          all_3d_model_predictions <- rbind(all_3d_model_predictions, 
                                            final_3d_over_filters)
        }
        
        # 2D results: merge absent and present results
        final_2d_over_filters <- merge.data.table(x = final_2d_over_filters[[1]], 
                                                  y = final_2d_over_filters[[2]], 
                                                  by = c("filter_id", "y", "x", "SF", "Ori"))
        final_2d_over_filters[ , sum_resp_diff := sum_resp_present - sum_resp_absent]
        # trial identifiers
        final_2d_over_filters[ , move_direction_f := move_direction_now]
        final_2d_over_filters[ , subj_id := subj_now]
        final_2d_over_filters[ , start_left_f := start_left_now]
        final_2d_over_filters[ , ID := ID_now]
        
        # aggregate 2D
        final_2d_over_filters_agg <- final_2d_over_filters[ , .(sum_resp_diff = mean(sum_resp_diff), 
                                                                max_resp_diff = max(sum_resp_diff), 
                                                                sum_resp_present = mean(sum_resp_present), 
                                                                max_resp_present = max(sum_resp_present), 
                                                                sum_resp_absent = mean(sum_resp_absent), 
                                                                max_resp_absent = max(sum_resp_absent)
        ), 
        by = .(filter_id, Ori, SF, 
               move_direction_f, subj_id, start_left_f, ID)]
        rm(final_2d_over_filters) # we don't need this anymore now
        
        # have a look, differences
        p_final_2d_over_filters_agg <- plot_grid(
          ggplot(data = final_2d_over_filters_agg, 
                 aes(x = Ori, y = SF, fill = sum_resp_diff)) + 
            geom_raster() + 
            coord_cartesian(expand = FALSE) + scale_y_log10() + 
            theme_minimal() + 
            # scale_fill_viridis_c(limits = c(min(final_2d_over_filters_agg[ , c("sum_resp_absent_diff", "sum_resp_present_diff")]), 
            #                                 max(final_2d_over_filters_agg[ , c("sum_resp_absent_diff", "sum_resp_present_diff")]))) + 
            scale_fill_viridis_c() + 
            labs(title = paste(cond_def, move_direction_now, sep = ", ")), 
          nrow = 1, align = "hv")
        print(p_final_2d_over_filters_agg)
        
        
        # save 2D data
        all_2d_model_predictions <- rbind(all_2d_model_predictions, 
                                          final_2d_over_filters_agg)
        
        
      } # end of iterate over move directions
    }
  }
}
save(all_2d_model_predictions, file = "all_2d_model_predictions.rda", compress = "xz")
asfd

# 1. normalize according to baseline activity to stimulus
all_2d_model_predictions[ , sum_resp_absent_norm := sum_resp_absent / max(sum_resp_absent), 
                          by = .(ID, Ori)]
all_2d_model_predictions[ , sum_resp_present_norm := sum_resp_present / max(sum_resp_absent), 
                          by = .(ID, Ori)]
all_2d_model_predictions[ , sum_resp_diff_norm := sum_resp_diff / max(sum_resp_absent), 
                          by = .(ID, Ori)]

all_2d_model_predictions_agg <- all_2d_model_predictions[ , .(sum_resp_diff = mean(sum_resp_diff_norm), 
                                                              sum_resp_present = mean(sum_resp_present_norm), 
                                                              sum_resp_absent = mean(sum_resp_absent_norm)), 
                                                          by = .(move_direction_f, start_left_f, 
                                                                 subj_id, Ori, SF)]

# overall activity
ggplot(data = all_2d_model_predictions_agg, aes(x = Ori, y = SF, fill = sum_resp_diff  )) + 
  geom_tile() + 
  scale_y_log10() + coord_cartesian() + 
  theme_minimal() + 
  facet_grid(move_direction_f+start_left_f~subj_id) + 
  scale_fill_viridis_c()



```
















































# Let the model output EEG data

```{r}
n_trials_per_cell <- 1

# run through all conditions
all_2d_model_predictions <- NULL
all_3d_model_predictions <- NULL
for (start_left_now in unique(all_saccades$start_left_f)) {
  for (move_direction_now in unique(all_saccades[move_direction_f!="static", move_direction_f]) ) {
    for (subj_now in unique(all_saccades$subj_id)[1]) {
      # get the corresponding target noise patch
      noise_patch_now <- all_noise_patches[[as.numeric(subj_now)]]
      # get a set of random trials in that condition
      sampled_IDs <- sample(unique(all_saccades[move_direction_f==move_direction_now & start_left_f==start_left_now & subj_id==subj_now, 
                                                ID]))[1:n_trials_per_cell]
      # run through those
      for (n_trial_now in 1:n_trials_per_cell) {
        cond_def <- paste(start_left_now, move_direction_now, subj_now, n_trial_now, sep = ", ")
        print(cond_def)
        # get trial data
        ID_now <- sampled_IDs[n_trial_now]
        trial_sac <- all_saccades[ID==ID_now]
        # remove samples were target is moving to simulate the absent condition
        trial_sac_absent <- copy(trial_sac)
        trial_sac_absent <- trial_sac_absent[stim_moving==FALSE]
        # check?
        #plot(trial_sac$time_stim_on, trial_sac$X)
        #points(trial_sac_absent$time_stim_on, trial_sac_absent$X, col = "red")
        
        # run model twice
        final_2d_over_filters <- vector(mode = "list", length = 2)
        final_3d_over_filters <- vector(mode = "list", length = 2)
        for (model_cond_i in 1:2) { # iterate over present and absent
          # in what condition?
          if (model_cond_i==1) { 
            model_cond <- "present"
            trial_sac_test <- trial_sac
          } else {
            model_cond <- "absent"
            trial_sac_test <- trial_sac_absent
          }
          # we'll test the moving condition (requiring sec sacc) relative to static condition (requiring no sac)
          staticmoving_2d_over_filters <- vector(mode = "list", length = 2)
          staticmoving_3d_over_filters <- vector(mode = "list", length = 2)
          for (staticmoving_i in 1:2) {
            # in moving or static condition?
            if (staticmoving_i==1) { 
              staticmoving_cond <- "static"
              retinal_x = trial_sac_test$retinal_x_static 
              retinal_y = trial_sac_test$retinal_y_static 
              retinal_t = trial_sac_test$time_sac_off 
            } else {
              staticmoving_cond <- "moving"
              retinal_x = trial_sac_test$retinal_x
              retinal_y = trial_sac_test$retinal_y
              retinal_t = trial_sac_test$time_sac_off 
            }
            # run model
            print(paste(model_cond, staticmoving_cond))
            v1_output <- v1(stim_mat = noise_patch_now, # the noise patch as a matrix
                            gabor_list = gabor_list_heiko, # a list of gabor filters created by 'get_gabor_filter_bank'
                            normalize_RFs = TRUE, 
                            irf_df = irf_space, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                            normalize_IRFs = TRUE, 
                            signal_x = retinal_x, 
                            signal_y = retinal_y, 
                            signal_t = retinal_t, 
                            output_full_sequences = TRUE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                            signal_x_range = c(min(c(trial_sac$retinal_x, trial_sac$retinal_x_static)), 
                                               max(c(trial_sac$retinal_x, trial_sac$retinal_x_static))), 
                            signal_y_range = c(min(c(trial_sac$retinal_y, trial_sac$retinal_y_static)), 
                                               max(c(trial_sac$retinal_y, trial_sac$retinal_y_static))), 
                            signal_t_range = c(min(trial_sac$time_sac_off), max(trial_sac$time_sac_off)), # the min and max of temporal scale
                            no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                            use_half_precision = TRUE, # if a GPU is available, use half-precision?
                            spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                            temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                            normalize_override = FALSE, 
                            use_normalization_pool = TRUE, # use normalization pool?
                            debug_mode = FALSE, # shows output at every step
                            show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                            final_maps_filename = paste0("v1_output_current_", model_cond, 
                                                         "_", staticmoving_cond, ".pdf") # name of the pdf file if show_final_maps==TRUE
            )
            # get Oris and SFs
            SFs_of_filters <- v1_output[[2]]
            Oris_of_filters <- v1_output[[3]]
            if (!is.null(v1_output[[4]])) {
              do_3d_stuff <- TRUE
            } else {
              do_3d_stuff <- FALSE
            }
            
            ## 1. summed-over-time model responses
            # create data table from summed output
            df_final_2d <- as.data.table(v1_output[[1]])
            colnames(df_final_2d) <- c("filter_id", "y", "x", 
                                       paste("sum_resp", model_cond, staticmoving_cond, sep = "_"))
            SF_col <- vector(mode = "numeric", length = nrow(df_final_2d))
            Ori_col <- vector(mode = "numeric", length = nrow(df_final_2d))
            for (r in 1:nrow(df_final_2d)) {
              SF_col[r] <- SFs_of_filters[df_final_2d$filter_id[r]]
              Ori_col[r] <- Oris_of_filters[df_final_2d$filter_id[r]]
            }
            df_final_2d[ , SF := SF_col]
            df_final_2d[ , Ori := Ori_col]
            # save:
            staticmoving_2d_over_filters[[staticmoving_i]] <- df_final_2d
            
            ## 2.1 reduce size of 3d matrix
            if (do_3d_stuff) {
              use_delayed_array <- "RleArray" # "HDF5Array" or "RleArray"
              if (use_delayed_array == "SparseArray") { package_name <- "SparseArray" }
              if (use_delayed_array == "RleArray") { package_name <- "DelayedArray" }
              if (use_delayed_array == "HDF5Array") { package_name <- "HDF5Array" }
              require(package_name, character.only = TRUE)
              print(paste("Compressing to", use_delayed_array, "..."))
              if (staticmoving_i==1) {
                v1_output_3d_static <- as(v1_output[[4]], use_delayed_array) 
                if (use_delayed_array != "SparseArray") { str(v1_output_3d_static) }
              } else if (staticmoving_i==2) {
                v1_output_3d_moving <- as(v1_output[[4]], use_delayed_array) 
                if (use_delayed_array != "SparseArray") { str(v1_output_3d_moving) }
              }
              print("Done.")
              # remove that massive list from memory
              rm(v1_output)
              gc() # ... and clean up
            }
            
            ## 2.2 model responses resolved over time 
              # TYPE 1: difference for each filter level, then abs-summed
              # TYPE 2: downsampled in space, then summed across all filters
            if (do_3d_stuff) {
              if (staticmoving_i==2) {
                print("Computing difference between moving and static from delayed arrays...")
                # compute difference
                are_equal(dim(v1_output_3d_static), dim(v1_output_3d_moving))
                if (use_delayed_array!="SparseArray") {
                  assert_that(is(v1_output_3d_moving, use_delayed_array) & is(v1_output_3d_static, use_delayed_array))
                  v1_output_3d_diff <- abs(v1_output_3d_moving - v1_output_3d_static) # these are all delayed operations
                  # if (!is(v1_output_3d_diff, use_delayed_array)) {
                  #   v1_output_3d_diff <- as(v1_output_3d_diff, use_delayed_array)
                  #   assert_that(is(v1_output_3d_diff, use_delayed_array))
                  # }
                  are_equal(dim(v1_output_3d_static), dim(v1_output_3d_diff))
                }
                # Parallelize here. Inspired by: https://www.glennklockwood.com/data-intensive/r/foreach-parallelism.html
                require(foreach)
                if (use_delayed_array!="SparseArray") {
                  n_cores <- length(unique(SFs_of_filters)) # use parallelization not for sparse arrays - too memory intense
                  require(doMC)
                  registerDoMC(length(unique(SFs_of_filters)))
                  `%dofun%` <- `%dopar%`
                } else {
                  `%dofun%` <- `%do%`
                }
                
                # run through filters (means and sums take a long time with RleArrays)
                df_diff_resp_over_time <- 
                  foreach(filter_i = 1:dim(v1_output_3d_static)[1], .verbose = FALSE, .errorhandling = "stop", 
                          .combine = 'rbind', 
                          .packages = c("assertthat", "data.table", package_name) ) %dofun% {
                            print(paste("Computing means ... ", filter_i, "of", dim(v1_output_3d_static)[1]))
                            # current SF and Orientation
                            SF_now <- SFs_of_filters[filter_i]
                            Ori_now <- Oris_of_filters[filter_i]
                            ## from delayed/sparse array get:
                            # differences
                            if (use_delayed_array!="SparseArray") { # it's a delayed operation anyway, no need to save memory
                              v1_output_3d_filter_diff_sums_over_time <- apply(X = v1_output_3d_diff[filter_i, , , ], 
                                                                               MARGIN = 3, 
                                                                               FUN = mean)
                            } else { # compute the difference right here to save memory
                              v1_output_3d_filter_diff_sums_over_time <- apply(X = abs(v1_output_3d_moving[filter_i, , , ] - 
                                                                                         v1_output_3d_static[filter_i, , , ]), 
                                                                               MARGIN = 3, 
                                                                               FUN = mean)
                            }
                            
                            # absolute means
                            v1_output_3d_filter_moving_sums_over_time <- apply(X = v1_output_3d_moving[filter_i, , , ], 
                                                                               MARGIN = 3,
                                                                               FUN = mean)
                            v1_output_3d_filter_static_sums_over_time <- apply(X = v1_output_3d_static[filter_i, , , ], 
                                                                               MARGIN = 3,
                                                                               FUN = mean)
                            are_equal(names(v1_output_3d_filter_moving_sums_over_time), names(v1_output_3d_filter_diff_sums_over_time))
                            are_equal(names(v1_output_3d_filter_moving_sums_over_time), names(v1_output_3d_filter_static_sums_over_time))
                            # combine
                            df_now <- data.table(SF = SF_now, Ori = Ori_now, 
                                                 time = as.numeric(names(v1_output_3d_filter_diff_sums_over_time)), 
                                                 diff_resp_sum = as.numeric(v1_output_3d_filter_diff_sums_over_time), 
                                                 moving_resp_sum = as.numeric(v1_output_3d_filter_moving_sums_over_time), 
                                                 static_resp_sum = as.numeric(v1_output_3d_filter_static_sums_over_time) )
                            colnames(df_now)[grepl(x = colnames(df_now), pattern = "_resp_sum")] <- 
                              paste(colnames(df_now)[grepl(x = colnames(df_now), pattern = "_resp_sum")], 
                                    model_cond, sep = "_")
                            # save
                            return(df_now)
                          } # end of iterate over filter_i
                # clean up
                if (use_delayed_array!="SparseArray") { rm(v1_output_3d_diff) }
                rm(v1_output_3d_moving, v1_output_3d_static)
                rm(SF_col, Ori_col, SFs_of_filters, Oris_of_filters)
                gc()
                print("Done.")
                # ... and save
                staticmoving_3d_over_filters[[staticmoving_i]] <- df_diff_resp_over_time
              }
            } # end of if (do_3d_stuff)
            
          } # now we have model outputs to both static and moving cases
          
          # this is static vs moving combined:
          staticmoving_2d_over_filters <- merge.data.table(x = staticmoving_2d_over_filters[[1]], 
                                                           y = staticmoving_2d_over_filters[[2]], 
                                                           by = c("filter_id", "y", "x", "SF", "Ori"))
          staticmoving_3d_over_filters <- rbindlist(staticmoving_3d_over_filters) # there is only one entry
          
          # save this
          final_2d_over_filters[[model_cond_i]] <- staticmoving_2d_over_filters
          final_3d_over_filters[[model_cond_i]] <- staticmoving_3d_over_filters
          
        } # now we have model outputs to both present and absent cases
        
        #asdf # this stops
        
        # merge absent and present results
        are_equal(sort(unique(final_3d_over_filters[[1]]$time)), sort(unique(final_3d_over_filters[[2]]$time)))
        final_3d_over_filters <- merge.data.table(x = final_3d_over_filters[[1]], 
                                                  y = final_3d_over_filters[[2]], 
                                                  by = c("SF", "Ori", "time"))
        final_3d_over_filters[ , move_direction_f := move_direction_now]
        final_3d_over_filters[ , subj_id := subj_now]
        final_3d_over_filters[ , start_left_f := start_left_now]
        final_3d_over_filters[ , ID := ID_now]
        
        # TO DO: plot output to present-static vs absent-static
        p_diff_3d_over_filters <- ggplot(data = final_3d_over_filters, 
                                           aes(x = time, y = diff_resp_sum_absent)) + 
          geom_line(aes(color = "1 absent")) + 
          geom_line(data = final_3d_over_filters, 
                    aes(x = time, y = diff_resp_sum_present, color = "2 present")) + 
          geom_line(data = final_3d_over_filters, 
                    aes(x = time, y = -1*(diff_resp_sum_present-diff_resp_sum_absent), color = "3 difference")) + 
          facet_grid(round(SF, 2)~round(Ori, 2)) + 
          theme_minimal() + 
          ggtitle(paste(cond_def, sep = ", ")) + 
          scale_color_viridis_d(end = 1)
        print(p_diff_3d_over_filters)
        
        # merge absent and present results
        final_2d_over_filters <- merge.data.table(x = final_2d_over_filters[[1]], 
                                                  y = final_2d_over_filters[[2]], 
                                                  by = c("filter_id", "y", "x", "SF", "Ori"))
        # compute the difference
        final_2d_over_filters[ , sum_resp_present_diff := abs(sum_resp_present_moving - sum_resp_present_static)]
        final_2d_over_filters[ , sum_resp_absent_diff := abs(sum_resp_absent_moving - sum_resp_absent_static)]
        # trial identifiers
        final_2d_over_filters[ , move_direction_f := move_direction_now]
        final_2d_over_filters[ , subj_id := subj_now]
        final_2d_over_filters[ , start_left_f := start_left_now]
        final_2d_over_filters[ , ID := ID_now]
        
        # aggregate 2D
        final_2d_over_filters_agg <- final_2d_over_filters[ , .(sum_resp_present_diff = mean(sum_resp_present_diff), 
                                                                sum_resp_absent_diff = mean(sum_resp_absent_diff), 
                                                                sum_resp_present_moving = mean(sum_resp_present_moving), 
                                                                sum_resp_present_static = max(sum_resp_present_static), 
                                                                sum_resp_absent_moving = mean(sum_resp_absent_moving), 
                                                                sum_resp_absent_static = max(sum_resp_absent_static)
                                                                ), 
                                                            by = .(filter_id, Ori, SF, 
                                                                   move_direction_f, subj_id, start_left_f, ID)]
        rm(final_2d_over_filters) # we don't need this anymore now
        
        # have a look, differences
        p_final_2d_over_filters_agg <- plot_grid(
          ggplot(data = final_2d_over_filters_agg, 
                 aes(x = Ori, y = SF, fill = sum_resp_present_diff)) + 
            geom_raster() + 
            coord_cartesian(expand = FALSE) + scale_y_log10() + 
            theme_minimal() + 
            scale_fill_viridis_c(limits = c(min(final_2d_over_filters_agg[ , c("sum_resp_absent_diff", "sum_resp_present_diff")]), 
                                            max(final_2d_over_filters_agg[ , c("sum_resp_absent_diff", "sum_resp_present_diff")]))) + 
            labs(title = paste(cond_def, "present", sep = ", ")), 
          ggplot(data = final_2d_over_filters_agg, 
                 aes(x = Ori, y = SF, fill = sum_resp_absent_diff)) + 
            geom_raster() + 
            coord_cartesian(expand = FALSE) + scale_y_log10() + 
            theme_minimal() + 
            scale_fill_viridis_c(limits = c(min(final_2d_over_filters_agg[ , c("sum_resp_absent_diff", "sum_resp_present_diff")]), 
                                            max(final_2d_over_filters_agg[ , c("sum_resp_absent_diff", "sum_resp_present_diff")]))) + 
            labs(title = paste(cond_def, "absent", sep = ", ")), 
          nrow = 1, align = "hv")
        print(p_final_2d_over_filters_agg)
        
        
        # save that data
        all_2d_model_predictions <- rbind(all_2d_model_predictions, 
                                          final_2d_over_filters_agg)
        all_3d_model_predictions <- rbind(all_3d_model_predictions, 
                                          final_2d_over_filters_agg)
        
      }
    }
  }
}
save(all_2d_model_predictions, file = "all_2d_model_predictions.rda", compress = "xz")
asfd

# 1. normalize according to baseline activity to stimulus
all_2d_model_predictions[ , sum_resp_absent_norm := sum_resp_absent / max(sum_resp_absent), 
                          by = .(ID, Ori)]
all_2d_model_predictions[ , sum_resp_present_norm := sum_resp_present / max(sum_resp_absent), 
                          by = .(ID, Ori)]
all_2d_model_predictions[ , sum_resp_diff_norm := sum_resp_diff / max(sum_resp_absent), 
                          by = .(ID, Ori)]

all_2d_model_predictions_agg <- all_2d_model_predictions[ , .(sum_resp_diff = mean(sum_resp_diff_norm), 
                                                              sum_resp_present = mean(sum_resp_present_norm), 
                                                              sum_resp_absent = mean(sum_resp_absent_norm)), 
                                                          by = .(move_direction_f, start_left_f, 
                                                                 subj_id, Ori, SF)]

# overall activity
ggplot(data = all_2d_model_predictions_agg, aes(x = Ori, y = SF, fill = sum_resp_diff  )) + 
  geom_tile() + 
  scale_y_log10() + coord_cartesian() + 
  theme_minimal() + 
  facet_grid(move_direction_f+start_left_f~subj_id) + 
  scale_fill_viridis_c()



```

# A few additional model tests

To evaluate the model, let's see whether it can make predictions about Kelly's stabilized motion surface...

```{r}
source("kelly_vel.R") # for Kelly's (1979) function
source("create_flicker_stim.R")
source("get_gabor_field_simple.R")

# globals
(test_SFs <- exp(seq(log(0.5), log(2.5), length.out = 3)))
(test_TFs <- exp(seq(log(0.5), log(50), length.out = 30)))
pres_dur <- 4 * 8 * 28 * (1000/120)
override_normalization <- TRUE
use_normal_test_RF <- FALSE # TRUE: linear Gabors, FALSE: log Gabors

# loop over SFs and TFs
prob_sums <- NULL
for (test_flicker in c(TRUE, FALSE)) {
  if (test_flicker) { what_now <- "flicker" } else { what_now <- "motion" }
  for (test_SF in test_SFs) {
    for (test_TF in test_TFs) {
      condition_now <- paste(what_now, round(test_SF, 2), round(test_TF, 2))
      print(condition_now)
      # both for flicker and motion:
      flicker_time <- seq(0, pres_dur, by = temporal_resolution)
      flicker_ramp <- dnorm(x = flicker_time, mean = pres_dur/2, 
                            sd = pres_dur/8) # SD OF GAUSSIAN RAMP! 1/8
      flicker_ramp <- flicker_ramp/max(abs(flicker_ramp)) # normalize to 1
      if (test_flicker) { # get flicker stimulus
        flicker_seq <- create_flicker_stim(dur = pres_dur, t_res = temporal_resolution, tfreq = test_TF)
        flicker_seq <- flicker_seq * flicker_ramp
        control_seq <- flicker_seq
        #plot(flicker_time, flicker_seq, type = "l", xlim = c(3000, 4000))
        #spec.ar(flicker_seq)
      } else { # get motion stimulus
        pcpf_motion = temporal_resolution/1000 * test_TF * 360
        phase_seq <- cumsum(rep(pcpf_motion, times = length(flicker_time)))
        motion_seq <- sin(deg2rad(phase_seq))*flicker_ramp
        control_seq <- motion_seq
        #plot(flicker_time, motion_seq, type = "l", xlim = c(3000, 4000))
        #spec.ar(motion_seq)
      }
      # get Gabor sequence according to the criteria 
      for (t_i in 1:length(flicker_time)) {
        if (test_flicker) {
          intensity_now <- flicker_seq[t_i]
          gabor_now <- get_gabor_field_simple(rf_freq_dva = test_SF, rf_width_dva = (1/test_SF) / 2, 
                                              rf_ori = 0, 
                                              rf_amp = intensity_now, 
                                              rf_phase_rad = pi/4, 
                                              scr.ppd = scr.ppd, ppd_scaler = 1/spatial_resolution, 
                                              mesh_size_scaler = 3, 
                                              create_aperture = TRUE, gaussian_aperture = TRUE)
        } else {
          intensity_now <- flicker_ramp[t_i]
          phase_now <- phase_seq[t_i]
          gabor_now <- get_gabor_field_simple(rf_freq_dva = test_SF, rf_width_dva = (1/test_SF) / 2, 
                                              rf_ori = 0, 
                                              rf_amp = intensity_now, 
                                              rf_phase_rad = deg2rad(phase_now), 
                                              scr.ppd = scr.ppd, ppd_scaler = 1/spatial_resolution, 
                                              mesh_size_scaler = 3, 
                                              create_aperture = TRUE, gaussian_aperture = TRUE)
        }
        if (t_i==1) {
          flicker_gabor_seq <- array(0, dim = c(dim(gabor_now[[1]]), length(flicker_time)), 
                                     dimnames = list(dimnames(gabor_now[[1]])[[1]], 
                                                     dimnames(gabor_now[[1]])[[2]], 
                                                     flicker_time))
        }
        flicker_gabor_seq[ , ,t_i] <- gabor_now[[1]]
      }
      #plot_heatmap(flicker_gabor_seq[ , , 4000], do_print = TRUE)
      #plot_heatmap(flicker_gabor_seq[ , , 4001], do_print = TRUE)
      
      # did this work?
      if (test_flicker) {
        flicker_seq_control <- apply(X = flicker_gabor_seq, 
                                   FUN = function(x) { diff(range(x)) / 2}, 
                                   MARGIN = c(3))
        #  plot(flicker_time, flicker_seq_control, type = "l)
        #  flicker_seq_control_peaks <- as.numeric(round(kmeans(which(flicker_seq_control>0.6), centers = 2)$centers))
        #  plot_heatmap(mat = flicker_gabor_seq[ , ,flicker_seq_control_peaks[1]], use_limits = c(-1, 1))
        #  plot_heatmap(mat = flicker_gabor_seq[ , ,flicker_seq_control_peaks[2]], use_limits = c(-1, 1))
      }
      
      # shall we test multiple SFs?
      test_SF_now <- test_SF #c(test_SF/1.4, test_SF, test_SF*1.4) # 1.4 according to Heiko's SF spacing
      # get receptive field/s
      if (use_normal_test_RF) { # normal Gabors
        test_RF <- get_gabor_filter_bank(
          all_SF = test_SF_now, 
          all_Ori = c(0), # positive means clockwise
          use_log_gabor_heiko = FALSE, 
          rf_width_override = NaN, # cpd, set to NaN to let the function determine RF field size
          scr_ppd = scr.ppd, 
          ppd_scaler = 1/spatial_resolution,
          make_gaussian_aperture = TRUE)
      } else { # log Gabors
        test_RF <- get_gabor_filter_bank(
          all_SF = test_SF_now, 
          all_Ori = c(0), # positive means clockwise
          use_log_gabor_heiko = TRUE, 
          rf_width_override = (1/test_SF_now) / 2 * 8, # let's assume 8 x RF SD
          scr_ppd = scr.ppd, 
          ppd_scaler = 1/spatial_resolution,
          make_gaussian_aperture = TRUE)
      }
      print(paste("dim(test_RF[[1]][[1]]) =", paste(dim(test_RF[[1]][[1]]), collapse = ",")))
      
      # ... and corresponding temporal response function
      test_IRF <- get_IRF_df(SFs = test_SF_now, 
                             irf_time_range = c(0, 200), 
                             temporal_resolution = temporal_resolution)
      # run our function
      test_RF_output <- v1(mat_over_t = flicker_gabor_seq,
                           gabor_list = test_RF, # a list of gabor filters created by 'get_gabor_filter_bank'
                           irf_df = test_IRF, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                           normalize_IRFs = TRUE, # ?????
                           norm_sigma = 0.07, 
                           normalize_override = override_normalization, # OVERRIDE NORMALIZATION???
                           use_normalization_pool = FALSE, 
                           output_full_sequences = TRUE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                           signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                           no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                           use_half_precision = FALSE, # if a GPU is available, use half-precision?
                           spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                           temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                           debug_mode = FALSE, # shows output at every step
                           show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                           final_maps_filename = "" # name of the pdf file if show_final_maps==TRUE
      )
      # get our full output
      summed_over_time_RF_output <- test_RF_output[[1]]
      plot_heatmap(summed_over_time_RF_output[1, , ])
      full_RF_output <- test_RF_output[[4]]
      full_RF_output <- apply(X = full_RF_output, 
                              FUN = function(x) { max(x) }, 
                              MARGIN = c(1, 4))
      dim(full_RF_output)
      apply(full_RF_output, FUN = function(x) {sum(x)}, MARGIN = c(1))
      #plot(flicker_time, full_RF_output, type = "l")
      
      # probability summation
      beta <- 1.2
      if (dim(test_RF_output[[1]])[1]==1) { # only one filter!
        prob_sum <- trapz(x = flicker_time, y = abs(full_RF_output)^beta )^(1/beta)
        simple_sum <- max(summed_over_time_RF_output[1, , ])
        simple_sum_control <- apply(full_RF_output, FUN = function(x) {sum(x)}, MARGIN = c(1))
      } else { # multiple filters
        prob_sum <- 0
        simple_sum <- 0
        simple_sum_control <- 0
        for (it in 1:dim(full_RF_output)[1]) {
          prob_sum <- prob_sum + trapz(x = flicker_time, y = abs(full_RF_output[it, ])^beta )^(1/beta)
          simple_sum <- simple_sum + max(summed_over_time_RF_output[it, , ])
          simple_sum_control <- simple_sum_control + sum(full_RF_output[it, ])
        }
      }
      (prob_sum_now <- data.table(motion_flicker = what_now, 
                                 SF = test_SF, TF = test_TF, 
                                 prob_sum = prob_sum, 
                                 simple_sum = simple_sum,
                                 simple_sum_control = simple_sum_control
      ))
      # prob_sums[motion_flicker=="motion" & SF==0.5]
      
      # add
      prob_sums <- rbind(prob_sums, prob_sum_now)
      
      
      ## UNDERSTAND: do the processing step by step
      do_understand <- FALSE
      if (do_understand) {
        require(smoothie)
        # step 1: spatial processing: convolution with Gabor filter
        filter_resp_zero <- apply(X = flicker_gabor_seq, 
                                  FUN = function(x) { max(kernel2dsmooth(x = x, K = test_RF[[1]][[1]])) }, 
                                  MARGIN = c(3))
        filter_resp_half <- apply(X = flicker_gabor_seq, 
                                  FUN = function(x) { max(kernel2dsmooth(x = x, K = test_RF[[1]][[2]])) }, 
                                  MARGIN = c(3))
        # step 2: convolution with temporal response
        irf_resp_zero <- zapsmall(convolve(filter_resp_zero, rev(test_IRF$irf), 
                                           type = "open"))[1:length(filter_resp_zero)]
        irf_resp_half <- zapsmall(convolve(filter_resp_half, rev(test_IRF$irf), 
                                           type = "open"))[1:length(filter_resp_zero)]
        # step 3: squaring
        irf_resp_squared <- sqrt(irf_resp_zero^2 + irf_resp_half^2)
        
        # plot this
        par(mfrow = c(2, 2))
        # stim
        plot(flicker_time, control_seq, type = "l", col = "black", main = condition_now)
        # gabor resp
        plot(flicker_time, filter_resp_zero, type = "l", col = "blue")
        lines(flicker_time, filter_resp_half, col = "red")
        # IRF resp
        plot(flicker_time, irf_resp_zero, type = "l", col = "blue")
        lines(flicker_time, irf_resp_half, col = "red")
        # squared resp
        plot(flicker_time, irf_resp_squared, type = "l")
        par(mfrow = c(1, 1))
      }
      
    }# end of loop over TFs
  } # end of loop over SFs
} # end of loop over flicker/motion

# make a kelly reference
prob_sums[ , kelly_pred := kelly_vel(SF, TF/SF)]
prob_sums[motion_flicker=="flicker", kelly_pred := kelly_pred / 2]

# this is our output
ggplot(prob_sums, aes(x = TF, y = simple_sum, color = SF, group = SF)) + 
  geom_line(data = prob_sums, aes(x = TF, y = kelly_pred, color = SF, group = SF), 
            linetype = "dotted", linewidth = 1.5) + 
  geom_line(linewidth = 1.5) + 
  scale_x_log10() + scale_y_log10(breaks = c(0.01, 0.1, 1, 10, 100), 
                                  labels = c(0.01, 0.1, 1, 10, 100)) + 
  annotation_logticks() + 
  scale_color_viridis_c(trans = "log10") + 
  theme_minimal() + 
  facet_wrap(~motion_flicker) + 
  labs(x = "Temporal frequency [Hz]", y = "Contrast sensitivity")
# check the ratios
prob_sums[motion_flicker=="motion", simple_sum] / prob_sums[motion_flicker=="flicker", simple_sum]
prob_sums[motion_flicker=="motion", prob_sum] / prob_sums[motion_flicker=="flicker", prob_sum]
prob_sums[motion_flicker=="motion", kelly_pred] / prob_sums[motion_flicker=="flicker", kelly_pred]


```

Try Anderson & Burr's function and have a look at the effect of envelope width on the Gabor frequency spectrum. We should see a widening of the frequency spectrum with a smaller Gabor aperture. 

```{r}
anderson_burr_gabor <- function(s, sigma, fk) {
  exp(-s^2 / (2*sigma^2)) * sin(2*pi*fk*s)
}

some_space <- seq(-10, 10, 0.1) # in degrees of visual angle
# compute some gabor
gabor_sf <- 0.5
some_gabor <- anderson_burr_gabor(s = some_space, sigma = 0.5, fk = gabor_sf)
some_gabor_2 <- anderson_burr_gabor(s = some_space, sigma = 1, fk = gabor_sf)
some_gabor_3 <- anderson_burr_gabor(s = some_space, sigma = 2, fk = gabor_sf)
# compute spectra of them
some_gabor_spec <- spectrum(some_gabor, plot = FALSE)
some_gabor_2_spec <- spectrum(some_gabor_2, plot = FALSE)
some_gabor_3_spec <- spectrum(some_gabor_3, plot = FALSE)
# plot both Gabors and spectrum
par(mfrow = c(1, 2))
plot(some_space, some_gabor, type = "l", main = "Gabor envelope", 
     ylim = c((min(c(some_gabor, some_gabor_2, some_gabor_3))), 
              (max(c(some_gabor, some_gabor_2, some_gabor_3)))))
lines(some_space, some_gabor_2, col = "red")
lines(some_space, some_gabor_3, col = "blue")
plot(some_gabor_spec$freq, log10(some_gabor_spec$spec), type = "l", main = "Gabor freq spectrum", 
     ylim = c(log10(min(c(some_gabor_spec$spec, some_gabor_2_spec$spec, some_gabor_3_spec$spec))), 
              log10(max(c(some_gabor_spec$spec, some_gabor_2_spec$spec, some_gabor_3_spec$spec)))) )
lines(some_gabor_2_spec$freq, log10(some_gabor_2_spec$spec), col = "red")
lines(some_gabor_3_spec$freq, log10(some_gabor_3_spec$spec), col = "blue")
par(mfrow = c(1, 1))
```

Indeed. 

Systematically investigate the role of Gabor stimulus SD and receptive field size. 
Ideally, 

```{r}
# first check whether the modified functions produce the same results as package smoothie
source("kernel2dsmooth_fftw.R")
source("Fourier2d_fftw.R")
require(smoothie)
test_mat <- randn(500, 500)
test_kernel <- get_gabor_field_simple(rf_freq_dva = 0.5, 
                                        rf_width_dva = 1, # aperture width would be 2*SD of the gaussian aperture
                                        rf_ori = 0, 
                                        rf_amp = 1, 
                                        rf_phase_rad = 0, 
                                        scr.ppd = 50, 
                                        ppd_scaler = 1, mesh_size_scaler = 3, 
                                        create_aperture = TRUE, gaussian_aperture = TRUE)[[1]]
test_mat_fft <- Fourier2d(test_mat, kdim = dim(test_kernel))
test_mat_fftw <- Fourier2d_fftw(test_mat, kdim = dim(test_kernel))
test_mat_fft[1:5, 1:5]
test_mat_fftw[1:5, 1:5]
test_conv_fft <- kernel2dsmooth(x = test_mat, X = test_mat_fft, K = test_kernel)
test_conv_fftw <- kernel2dsmooth_fftw(x = test_mat, X = test_mat_fftw, K = test_kernel)
plot_grid(plot_heatmap(test_conv_fft), plot_heatmap(test_conv_fftw))
rm(test_mat, test_kernel, test_mat_fft, test_mat_fftw, test_conv_fft, test_conv_fftw)

# load aux functions:
source("normSum.R")
source("get_fft2.R")

# define the parameters we want to test 
test_SFs <- exp(seq(log(0.5), log(10), length.out = 9))
test_SD_facs <- exp(seq(log(0.1), log(10), length.out = 9)) # relative to 1/test_SFs
sd_test_scr.ppd <- 100
max_test_SD <- (1/min(test_SFs))/2 * max(test_SD_facs)
max_field_size <- round(max_test_SD * 2 * 3 * sd_test_scr.ppd)+10 # SD * 2 (=aperture width) * 3 (mesh_size_scaler)
if ( mod(max_field_size, 2)==0 ) { max_field_size <- max_field_size + 1}
max_field_center <- ceiling(max_field_size / 2)


# run through spatial frequencies and different Gabor SDs
test_SD_res <- NULL

for (test_SF in test_SFs) {
  for (test_SD_fac in test_SD_facs) {
    condition_def <- paste0("SF = ", round(test_SF, 2), ", SD_fac = ", round(test_SD_fac, 2))
    print(condition_def)
    # determine aperture size
    base_test_SD <- (1/test_SF) / 2 
    test_SD <- base_test_SD * test_SD_fac
    ## create stimulus
    gabor_now <- get_gabor_field_simple(rf_freq_dva = test_SF, 
                                        rf_width_dva = 2*test_SD, # aperture width would be 2*SD of the gaussian aperture
                                        rf_ori = 0, 
                                        rf_amp = 1, 
                                        rf_phase_rad = pi/4, 
                                        scr.ppd = sd_test_scr.ppd, 
                                        ppd_scaler = 1/spatial_resolution, 
                                        mesh_size_scaler = 3, 
                                        create_aperture = TRUE, gaussian_aperture = TRUE)
    print(paste("dim(gabor_now[[1]]) =", paste(dim(gabor_now[[1]]), collapse = ",")))
    # pad so that we achieve a fixed size
    stim_now <- matrix(0, nrow = max_field_size, ncol = max_field_size)
    stim_now[seq(round(max_field_center-nrow(gabor_now[[1]])/2), 
                 round(max_field_center-nrow(gabor_now[[1]])/2)+nrow(gabor_now[[1]])-1),
             seq(round(max_field_center-ncol(gabor_now[[1]])/2),
                 round(max_field_center-ncol(gabor_now[[1]])/2)+ncol(gabor_now[[1]])-1)] <- gabor_now[[1]]
    ## create receptive field
    use_normal_test_RF <- FALSE
    if (use_normal_test_RF) { # normal Gabors
      test_RF <- get_gabor_filter_bank(
        all_SF = test_SF, 
        all_Ori = c(0), # positive means clockwise
        use_log_gabor_heiko = FALSE, 
        rf_width_override = (1/test_SF) , # cpd, set to NaN to let the function determine RF field size
        scr_ppd = sd_test_scr.ppd, 
        ppd_scaler = 1/spatial_resolution,
        make_gaussian_aperture = TRUE)
    } else { # log Gabors
      test_RF <- get_gabor_filter_bank(
        all_SF = test_SF, 
        all_Ori = c(0), # positive means clockwise
        use_log_gabor_heiko = TRUE, 
        rf_width_override = (1/test_SF) / 2 * 8,
        scr_ppd = sd_test_scr.ppd, 
        ppd_scaler = 1/spatial_resolution,
        make_gaussian_aperture = TRUE)
    }
    print(paste("dim(test_RF[[1]][[1]]) =", paste(dim(test_RF[[1]][[1]]), collapse = ",")))
    
    ## convolute the stimulus with the RF
    stim_now_fft <- Fourier2d_fftw(x = stim_now, kdim = dim(test_RF[[1]][[1]])) # to save time, compute fft of the stim here
    stim_conv_RF_1 <- kernel2dsmooth_fftw(x = stim_now, X = stim_now_fft, K = normSum(test_RF[[1]][[1]]) )
    stim_conv_RF_2 <- kernel2dsmooth_fftw(x = stim_now, X = stim_now_fft, K = normSum(test_RF[[1]][[2]]) )
    stim_conv_RF <- sqrt(stim_conv_RF_1^2 + stim_conv_RF_2^2)
    rm(stim_now_fft, stim_conv_RF_1, stim_conv_RF_2) # release memory
    
    ## analyze spectrum
    do_fft <- FALSE
    if (do_fft) {
      stim_now_spec <- get_fft2(stim_now, scr.ppd = sd_test_scr.ppd, downsample_factor = 10, 
                                use_torch = TRUE)
      # ggplot(data = stim_now_spec[[1]], aes(x = rf_ori, y = rf_freq_dva, color = power)) + 
      #   geom_point(alpha = 0.1) + 
      #   coord_cartesian(expand = FALSE) + theme_minimal() + 
      #   scale_color_viridis_c()
      p_stim_now_spec <- plot_heatmap(stim_now_spec[[4]], use_viridis = TRUE, do_print = FALSE) + ggtitle(condition_def)
      print(p_stim_now_spec)
      rm(stim_now_spec, p_stim_now_spec)
    }
    
    # readout summary values
    (stim_conv_readout <- data.table(SF = test_SF, SD = test_SD, 
                                     base_SD = base_test_SD, test_SD_fac = test_SD_fac, 
                                     conv_max = max(abs(stim_conv_RF)), 
                                     conv_sum = sum(abs(stim_conv_RF))) )
    rm(stim_conv_RF) # release memory
    print(stim_conv_readout)
    test_SD_res <- rbind(test_SD_res, stim_conv_readout)
    gc() # garbage cleanup
    
  }
}

# plot maximum value:
# 1. over aperture SDs
ggplot(data = test_SD_res, 
       aes(x = SD, y = conv_max, color = SF, group = SF)) + 
  geom_line() + 
  geom_vline(aes(xintercept = 1/SF, color = SF), linetype = "dotted") + 
  scale_x_log10() + scale_y_log10() + scale_color_viridis_c(trans = "log10") + 
  theme_minimal() + 
  labs(x = "Aperture SD", y = "Max of image convoluted with Gabor filter", color = "SF [cpd]")
# 2. over relative aperture SD
ggplot(data = test_SD_res, 
       aes(x = test_SD_fac, y = conv_max, color = SF, group = SF)) + 
  geom_line() + 
  geom_vline(aes(xintercept = 1, color = SF), linetype = "dotted") + 
  scale_x_log10() + scale_y_log10() + scale_color_viridis_c(trans = "log10") + 
  theme_minimal() + 
  labs(x = "Aperture SD relative to 1/SF/2", y = "Max of image convoluted with Gabor filter", color = "SF [cpd]")

```

Determine the spectrum of the RFs described by Anderson & Burr (1987)

```{r}
anderson_burr_RF_width <- function(SF) {
  widths <- rep(NaN, length(SF))
  widths[SF<1] <- exp(-0.5 * log(SF[SF<1]) + log(1))
  widths[SF>=1] <- 1/SF[SF>=1]
  return(widths)
}
# use Anderson & Burr's function
anderson_burr_df <- data.table(sf = exp(seq(log(0.01), log(10), length.out=10)))
anderson_burr_df$widths <- anderson_burr_RF_width(anderson_burr_df$sf)
# ... and plot results. Compare to their Figure 4
plot_grid(
  ggplot(data = anderson_burr_df, aes(x = sf, y = widths)) + 
    geom_line() + 
    scale_x_log10() + scale_y_log10() + 
    theme_minimal() + 
    labs(x = "SF [cpd]", y = "RF width in dva"), 
  ggplot(data = anderson_burr_df, aes(x = sf, y = widths/(1/sf))) + 
    geom_line() + 
    scale_x_log10() + scale_y_log10() + 
    theme_minimal() + 
    labs(x = "SF [cpd]", y = "RF width in number of cycles")
)

# run through all 
for (test_SF_i in 1:nrow(anderson_burr_df)) {
  condition_def <- paste0("SF=", round(anderson_burr_df$sf[test_SF_i], 2), 
                          ", width=", round(anderson_burr_df$widths[test_SF_i], 2))
  # create the stimulus
  gabor_now <- get_gabor_field_simple(rf_freq_dva = anderson_burr_df$sf[test_SF_i], 
                                      rf_width_dva = anderson_burr_df$widths[test_SF_i], 
                                      rf_ori = 0, 
                                      rf_amp = 1, 
                                      rf_phase_rad = pi/2, 
                                      scr.ppd = sd_test_scr.ppd, 
                                      ppd_scaler = 1/spatial_resolution, 
                                      mesh_size_scaler = 3, 
                                      create_aperture = TRUE, gaussian_aperture = TRUE)
  print(paste("dim(gabor_now[[1]]) =", paste(dim(gabor_now[[1]]), collapse = ",")))
  if (test_SF_i == 1) {
    max_field_size <- unique(dim(gabor_now[[1]]))+10
    if ( mod(max_field_size, 2)==0 ) { max_field_size <- max_field_size + 1}
    max_field_center <- ceiling(max_field_size / 2)
  }
  # pad so that we achieve a fixed size
  stim_now <- matrix(0, nrow = max_field_size, ncol = max_field_size)
  stim_now[seq(round(max_field_center-nrow(gabor_now[[1]])/2), 
               round(max_field_center-nrow(gabor_now[[1]])/2)+nrow(gabor_now[[1]])-1),
           seq(round(max_field_center-ncol(gabor_now[[1]])/2),
               round(max_field_center-ncol(gabor_now[[1]])/2)+ncol(gabor_now[[1]])-1)] <- gabor_now[[1]]
  print(paste("dim(stim_now) =", paste(dim(stim_now), collapse = ",")))
  # perform the FFT2
  stim_now_spec <- get_fft2(stim_now, scr.ppd = sd_test_scr.ppd, downsample_factor = 1, 
                                use_torch = TRUE)
  plot(stim_now_spec[[4]][round(nrow(stim_now_spec[[4]])/2), round(ncol(stim_now_spec[[4]])/2-10):round(ncol(stim_now_spec[[4]])/2+10)], 
       type = "l", main = condition_def)
  # ... and show:
  p_stim_now_spec <- plot_heatmap(stim_now_spec[[4]], use_viridis = TRUE, do_print = FALSE) + 
    ggtitle(condition_def)
  #spectrum(stim_now[round(nrow(stim_now)/2),], type = "l")
  
}




```


Test the effect of ocular drift on sensitivity, see Kelly, 1979 (Part I), Fig. 6. 

```{r}
# get the self-avoiding random walk function
source("self_avoiding_walk.R")
source("vecvel.R")

# produce a random walk, assuming the presentation used above
walk_pres_dur <- 4 * 8 * 28 * (1000/120) 
walk_scr.ppd <- 40
walk_start <- 50
walk_override_normalization <- TRUE
walk_sampling_rate <- 60
walk_time <- seq(0, walk_pres_dur, by = 1000/walk_sampling_rate)
walked_smooth <- self_avoiding_walk(n_steps = length(walk_time),  
                                    relax_rate = 0.001, 
                                    L = walk_start*2 + 1, 
                                    U_slope = 0.2, 
                                    U_m_sd=c(0.5, 0.2), 
                                    use_only_direct_neighbors = TRUE, 
                                    do_plot = TRUE, 
                                    do_final_smooth = TRUE)
# and check it's velocity, so that it approximately matches empirical drift speed
# see Drift speed distribution for one observer during normal head-free viewing (Aytekin, Victor & Rucci 2014)
walked_smooth_vel <- vecvel(as.matrix(walked_smooth/(walk_scr.ppd)), # convert to arcmin per second
                            SAMPLING = walk_sampling_rate) # the sampling rate determines the speed
hist(sqrt(walked_smooth_vel[ , 1]^2 + walked_smooth_vel[ , 2]^2))

# upsample drift motion
walk_x_approx <- approxfun(walk_time, walked_smooth[ ,1])
walk_y_approx <- approxfun(walk_time, walked_smooth[ ,2])
walk_time_1440Hz <- seq(0, walk_pres_dur, temporal_resolution)
walked_smooth_1440Hz <- cbind(walk_x_approx(walk_time_1440Hz), 
                              walk_y_approx(walk_time_1440Hz))
plot(walked_smooth[ ,1], walked_smooth[ ,2])
lines(walked_smooth_1440Hz[ ,1], walked_smooth_1440Hz[ ,2])

# what is the extent of the walk?
(walk_extent <- max(c(ceiling(diff(range(walked_smooth_1440Hz[,2]))), 
                      ceiling(diff(range(walked_smooth_1440Hz[,1]))))))

# test different SFs with our without stabilization
drift_gains <- c(0.0)
test_SFs <- exp(seq(log(0.5), log(10), length.out = 9))
test_VELs <- c(0.001)#, 0.012, 0.15, 3, 11)  # according to  Kelly (1979b) Figure 6
walk_use_IRFs <- FALSE

stab_res <- NULL

for (drift_gain in drift_gains) {
  for (test_VEL in test_VELs) {
    for (test_SF in test_SFs) {
      # compute TF
      test_TF = test_VEL * test_SF
      print(paste("drift =", drift_gain, ", SF =", round(test_SF, 2), ", VEL =", round(test_VEL, 5), ", TF =", round(test_TF, 5)))
      
      ## 1. create stimulus
      # both for flicker and motion:
      flicker_time <- seq(0, walk_pres_dur, by = temporal_resolution)
      flicker_ramp <- dnorm(x = flicker_time, mean = walk_pres_dur/2, 
                            sd = walk_pres_dur/8) # SD OF GAUSSIAN RAMP! 1/8
      flicker_ramp <- flicker_ramp/max(abs(flicker_ramp)) # normalize to 1
      test_flicker <- FALSE
      if (test_flicker) { # get flicker stimulus
        flicker_seq <- create_flicker_stim(dur = walk_pres_dur, t_res = temporal_resolution, tfreq = test_TF)
        flicker_seq <- flicker_seq * flicker_ramp
        control_seq <- flicker_seq
        #plot(flicker_time, flicker_seq, type = "l")
        #spec.ar(flicker_seq)
      } else { # get motion stimulus
        pcpf_motion = temporal_resolution/1000 * test_TF * 360
        phase_seq <- cumsum(rep(pcpf_motion, times = length(flicker_time)))
        motion_seq <- cos(deg2rad(phase_seq))*flicker_ramp
        control_seq <- motion_seq
        #plot(flicker_time, motion_seq, type = "l")
        #spec.ar(motion_seq)
      }
      # get Gabor sequence according to the criteria 
      for (t_i in 1:length(flicker_time)) {
        if (test_flicker) { # flicker:
          intensity_now <- flicker_seq[t_i]
          gabor_now <- get_gabor_field_simple(rf_freq_dva = test_SF, rf_width_dva = (1/min(test_SFs)), 
                                              rf_ori = 0, 
                                              rf_amp = intensity_now, 
                                              rf_phase_rad = pi/4, 
                                              scr.ppd = walk_scr.ppd, ppd_scaler = 1/spatial_resolution, 
                                              mesh_size_scaler = 4, 
                                              create_aperture = FALSE, # NO APERTURE!
                                              gaussian_aperture = TRUE)
        } else { # motion:
          intensity_now <- 1#flicker_ramp[t_i]
          phase_now <- phase_seq[t_i]
          gabor_now <- get_gabor_field_simple(rf_freq_dva = test_SF, rf_width_dva = (1/min(test_SFs)), 
                                              rf_ori = 0, 
                                              rf_amp = intensity_now, 
                                              rf_phase_rad = deg2rad(phase_now), 
                                              scr.ppd = walk_scr.ppd, ppd_scaler = 1/spatial_resolution, 
                                              mesh_size_scaler = 4, 
                                              create_aperture = FALSE, # NO APERTURE!
                                              gaussian_aperture = TRUE)
        }
        if (t_i==1) {
          target_view_y_pix <- round((1/test_SF) * 2 * walk_scr.ppd)
          target_view_x_pix <- round((1/test_SF) * 2 * walk_scr.ppd)
          flicker_gabor_seq <- array(0, dim = c(target_view_y_pix, 
                                                target_view_x_pix, 
                                                length(flicker_time)))
          padded_gabor_space <- matrix(0, nrow = nrow(gabor_now[[1]])+walk_extent*2, 
                                       ncol = ncol(gabor_now[[1]])+walk_extent*2)
          flicker_gabor_seq_center <- unique(dim(padded_gabor_space)[1:2]%/%2)
        }
        # embed the gabor in the center of this:
        padded_gabor_space[seq(round(flicker_gabor_seq_center-nrow(gabor_now[[1]])/2), 
                               round(flicker_gabor_seq_center-nrow(gabor_now[[1]])/2)+nrow(gabor_now[[1]])-1),
                           seq(round(flicker_gabor_seq_center-ncol(gabor_now[[1]])/2),
                               round(flicker_gabor_seq_center-ncol(gabor_now[[1]])/2)+ncol(gabor_now[[1]])-1)] <- 
          gabor_now[[1]]
        #    plot_heatmap(padded_gabor_space)
        # we want to interpolate from this image
        require(fields)
        image_obj <- list(x = 1:nrow(padded_gabor_space), y = 1:ncol(padded_gabor_space), 
                          z = padded_gabor_space) # here we subtract the midpoint
        # what is the current eye position with respect to the image?
        current_eye_x <- (walked_smooth_1440Hz[t_i, 1] - walk_start)*drift_gain + flicker_gabor_seq_center
        current_eye_y <- (walked_smooth_1440Hz[t_i, 2] - walk_start)*drift_gain + flicker_gabor_seq_center
        # get the current retinal view
        select_col <- seq(from = current_eye_x-target_view_x_pix/2, 
                          to = current_eye_x-target_view_x_pix/2+target_view_x_pix-1, 
                          by = 1 )
        assert_that(length(select_col)==target_view_y_pix )
        select_row <- seq(from = current_eye_y-target_view_y_pix/2, 
                          to = current_eye_y-target_view_y_pix/2+target_view_y_pix-1, 
                          by = 1 )
        assert_that(length(select_row)==target_view_y_pix )
        assert_that(length(select_row)==length(select_col))
        query_points <- expand.grid(rows = select_row, cols = select_col)
        # select current subimage via linear interpolation (requires package fields)
        target_image <- interp.surface( obj = image_obj, loc = query_points )
        target_image <- matrix(data = target_image, nrow = length(select_row), ncol = length(select_col), 
                               byrow = FALSE, dimnames = list(select_row, select_col))
        # convolve with a Gaussian envelope
        meshlist <- meshgrid(x = (1:ncol(target_image))-ncol(target_image)/2, 
                             y = (1:nrow(target_image))-nrow(target_image)/2) 
        target_gaussian_sd <- ((1/test_SF) / 2) * walk_scr.ppd
        target_gaussian <- exp(-(meshlist$X^2+meshlist$Y^2)/(2*target_gaussian_sd^2)) # compute gaussian aperture of stimulus
        target_image <- target_image * target_gaussian
        # put into frame
        flicker_gabor_seq[ , , t_i] <- target_image
      }
      print(paste("dim(flicker_gabor_seq) =", paste(dim(flicker_gabor_seq), collapse = ",")))
      # # see example
      # for (it in seq(1, 1000, 100)) {
      #   plot_heatmap(flicker_gabor_seq[ , ,4000+it], do_print = TRUE)
      # }

      ## RUN THE MODEL
      # shall we test multiple SFs?
      test_SF_now <- test_SF #c(test_SF/1.4, test_SF, test_SF*1.4) # 1.4 according to Heiko's SF spacing
      # get receptive field/s
      use_normal_test_RF <- TRUE
      if (use_normal_test_RF) { # normal Gabors
        test_RF <- get_gabor_filter_bank(
          all_SF = test_SF_now, 
          all_Ori = c(0), # positive means clockwise
          use_log_gabor_heiko = FALSE, 
          rf_width_override = NaN , # cpd, set to NaN to let the function determine RF field size
          scr_ppd = walk_scr.ppd, 
          ppd_scaler = 1/spatial_resolution,
          make_gaussian_aperture = TRUE)
      } else { # log Gabors
        test_RF <- get_gabor_filter_bank(
          all_SF = test_SF_now, 
          all_Ori = c(0), # positive means clockwise
          use_log_gabor_heiko = TRUE, 
          rf_width_override = (1/test_SF) / 2 * 8,
          scr_ppd = walk_scr.ppd, 
          ppd_scaler = 1/spatial_resolution,
          make_gaussian_aperture = TRUE)
      }
      print(paste("dim(test_RF[[1]][[1]]) =", paste(dim(test_RF[[1]][[1]]), collapse = ",")))
      
      # ... and corresponding temporal response function
      test_IRF <- get_IRF_df(SFs = test_SF_now, 
                             irf_time_range = c(0, 200), 
                             temporal_resolution = temporal_resolution)
      
      # run our function - WITH DRIFT
      if (walk_use_IRFs) {
        test_RF_output <- v1(mat_over_t = flicker_gabor_seq,
                             gabor_list = test_RF, # a list of gabor filters created by 'get_gabor_filter_bank'
                             irf_df = test_IRF, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                             norm_sigma = 0.07, 
                             normalize_override = walk_override_normalization, # OVERRIDE NORMALIZATION???
                             use_normalization_pool = FALSE, 
                             output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                             signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                             no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                             use_half_precision = FALSE, # if a GPU is available, use half-precision?
                             spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                             temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                             debug_mode = FALSE, # shows output at every step
                             show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                             final_maps_filename = "" # name of the pdf file if show_final_maps==TRUE
        )
      } else {
        test_RF_output <- v1(mat_over_t = flicker_gabor_seq,
                             gabor_list = test_RF, # a list of gabor filters created by 'get_gabor_filter_bank'
                             irf_df = NULL, # a data.frame of IRFs resolved in SF x time created by 'get_IRF_df'
                             norm_sigma = 0.07, 
                             normalize_override = walk_override_normalization, # OVERRIDE NORMALIZATION???
                             use_normalization_pool = FALSE, 
                             output_full_sequences = FALSE, # TRUE: returns complete sequences. WARNING: this is memory-intense
                             signal_x_range = NULL, signal_y_range = NULL, signal_t_range = NULL, # the min and max of temporal scale
                             no_CUDA = FALSE, # set to TRUE if you want to use CPU instead of GPU (but torch)
                             use_half_precision = FALSE, # if a GPU is available, use half-precision?
                             spatial_resolution = spatial_resolution, # spatial resolution of processing function in dva
                             temporal_resolution = temporal_resolution, # temporal resolution of processing function in milliseconds
                             debug_mode = FALSE, # shows output at every step
                             show_final_maps = TRUE,  # shows all resulting 2D maps at the end
                             final_maps_filename = "" # name of the pdf file if show_final_maps==TRUE
        )
      }
      
      
      # get our full output
      summed_over_time_RF_output <- test_RF_output[[1]]
      #plot_grid(plot_heatmap(summed_over_time_RF_output[1, , ]))
      
      # summation
      (prob_sum_now <- data.table(motion_flicker = what_now, 
                                  SF = test_SF, VEL = test_VEL, TF = test_TF, 
                                  drift_gain = drift_gain, 
                                  simple_sum = max(summed_over_time_RF_output[1, , ])
      ))
      stab_res <- rbind(stab_res, prob_sum_now)
    }
  }
}
stab_res
stab_res[ , kelly_sens := kelly_vel(sf = SF, v = VEL)]

# plot:
ggplot(stab_res, aes(x = SF, y = simple_sum/1000, color = VEL, group = VEL)) + 
  geom_line(linewidth = 1.5) + 
  geom_point(data = stab_res, aes(x = SF, y = kelly_sens, color = VEL, group = VEL), 
            size = 1.5) + 
  scale_x_log10() + scale_y_log10(limits = c(1, 300)) +  # limits = c(2, 500)
  annotation_logticks() + 
  scale_color_viridis_c(trans = "log10") + 
  theme_minimal() + 
  facet_wrap(~motion_flicker+drift_gain) + 
  labs(x = "Spatial frequency [cpd]", y = "Contrast sensitivity", color = "Velocity")

```

This is what the prediction should look like:
```{r}

kelly_fig_6 <- data.table(expand.grid(VEL = c(0.0001, 0.012, 0.15, 3, 11, 32), 
                                      SF = exp(seq(log(0.2), log(10), length.out = 30))))
kelly_fig_6[ , sens := kelly_vel(sf = SF, v = VEL)]

ggplot(kelly_fig_6, aes(x = SF, y = sens, color = as.factor(VEL) )) + 
  geom_line(linewidth = 1.5) + 
  scale_x_log10() + scale_y_log10(limits = c(0.1, 500)) + 
  annotation_logticks() + 
  scale_color_viridis_d() + 
  theme_minimal() + 
  labs(x = "Spatial frequency [cpd]", y = "Contrast sensitivity", color = "Velocity")
```

Try a fft-based temporal convolution:
```{r}
# test these parameters
test_SF <- 5
test_VEL <- 0.1
test_fft_pad <- 1001

# compute TF
test_TF = test_VEL * test_SF
print(paste("SF =", round(test_SF, 2), ", VEL =", round(test_VEL, 4), ", TF =", round(test_TF, 4)))

# create motion sequence:
flicker_time <- seq(0, walk_pres_dur, by = temporal_resolution)
flicker_ramp <- dnorm(x = flicker_time, mean = walk_pres_dur/2, 
                      sd = walk_pres_dur/8) # SD OF GAUSSIAN RAMP! 1/8
flicker_ramp <- flicker_ramp/max(abs(flicker_ramp)) # normalize to 1
test_flicker <- FALSE
pcpf_motion = temporal_resolution/1000 * test_TF * 360
phase_seq <- cumsum(rep(pcpf_motion, times = length(flicker_time)))
# get Gabor sequence according to the criteria 
for (t_i in 1:length(flicker_time)) {
    intensity_now <- 1#flicker_ramp[t_i]
    phase_now <- phase_seq[t_i]
    gabor_now <- get_gabor_field_simple(rf_freq_dva = test_SF, rf_width_dva = (1/test_SF), 
                                        rf_ori = 0, 
                                        rf_amp = intensity_now, 
                                        rf_phase_rad = deg2rad(phase_now), 
                                        scr.ppd = walk_scr.ppd, ppd_scaler = 1/spatial_resolution, 
                                        mesh_size_scaler = 2, 
                                        create_aperture = FALSE, # NO APERTURE!
                                        gaussian_aperture = TRUE)
    if (t_i==1) {
      flicker_gabor_seq <- array(0, dim = c(dim(gabor_now[[1]]), length(flicker_time)), 
                                 dimnames = list(dimnames(gabor_now[[1]])[[1]], 
                                                 dimnames(gabor_now[[1]])[[2]], 
                                                 flicker_time))
    }
    flicker_gabor_seq[ , ,t_i] <- gabor_now[[1]]
}
dim(flicker_gabor_seq)

# make it a tensor
test_mat <- torch_tensor(flicker_gabor_seq)
test_mat <- test_mat$permute(c(3, 1, 2))$unsqueeze(1)
dim(test_mat)
# temporal resp fun
resp_fun <- torch_tensor(test_IRF$irf)
resp_fun <- resp_fun$expand(c(test_mat$size(4), 1, resp_fun$size())) # expand to match matrix dimensions (x dimension)
resp_fun <- torch_cat(list(torch_zeros(c(resp_fun$size(1), resp_fun$size(2), resp_fun$size(3)-1)), # create the zero pad to center the TRF
                           resp_fun), dim = 3)
dim(resp_fun)
# reshape again
test_mat <- test_mat$permute(c(1, 3, 4, 2))$squeeze() # new dim: y, x, t, as initially
# perform standard conv
temporal_response_1 <- torch_conv1d(input = test_mat, 
                                    weight = resp_fun$flip(3) / max(resp_fun), 
                                    padding = resp_fun$size(3)%/%2, 
                                    groups = test_mat$size(2) # is the x dimension
)
# # did this work plots
# plot(test_mat[round(nrow(test_mat)/2), round(ncol(test_mat)/2), ])
# plot(abs(fftshift(fft(as_array(test_mat[round(nrow(test_mat)/2), round(ncol(test_mat)/2), ]) ))), type = "l")
# plot(resp_fun)
# plot(temporal_response_1[round(nrow(temporal_response_1)/2), round(ncol(temporal_response_1)/2), ])

# use the FFT version of the convolution
source("temporal_conv_IRF_fft.R")
temporal_response_2 <- temporal_conv_IRF_fft(test_mat = test_mat, 
                                             IRF = test_IRF$irf, IRF_time = test_IRF$irf_time, 
                                             pad = test_fft_pad)

# use Kelly's function directly
source("temporal_conv_kelly_fft.R")
temporal_response_3 <- temporal_conv_kelly_fft(test_mat = test_mat, SF = test_SF, 
                                               pad = test_fft_pad)

# plot all next to each other
par(mfrow=c(2, 2))
plot(temporal_response_1[round(nrow(temporal_response_1)/2), round(ncol(temporal_response_1)/2), ], 
     main = "standard IRF convolution")
plot(temporal_response_2[round(nrow(temporal_response_2)/2), round(ncol(temporal_response_2)/2), ], 
     main = "fft-based IRF convolution")
plot(temporal_response_3[round(nrow(temporal_response_3)/2), round(ncol(temporal_response_3)/2), ], 
     main = "fft-based Kelly convolution")
par(mfrow=c(1, 1))


```


